---
title: "Calibrate ET"
author: "Beni Stocker"
date: "`r Sys.Date()`"
# output:
#   html_document:
#     toc: true
#     toc_float: true
#     toc_depth: 4
#     number_sections: true
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
header-includes:
   - \usepackage{amsmath}
bibliography: bibliography.bib
---

## Rationale

```{r setup, include=FALSE}
library(rsofun)
load_dependencies_rsofun()
```

ET is simulated for the MCT-zroot study as:
$$
ET = a\; \text{fAPAR} \; \text{PET} \; G_s(D)
$$

Where the function $G_s(D)$ is taken from Oren et al. (2001) as
$$
G_s(D) = b - m \ln(D)
$$
This requires three additional model parameters for SPLASH/SOFUN: $a, b, m$.

```{r}
a <- 1
b <- 70
m <- 35

# from calibration, using RMSE:
a <- 0.1
b <- 3.841307
m <- 1.368163

# from calibration, using chi-squared:
a <- 0.100000        
b <- 2.444468        
m <- 0.010000 

# from calibration, using RMSE
a <- 0.9420240       
b <- 0.4974496       
m <- 0.0001000 

a <- 0.5       
b <- 0.4    
m <- 1

calc_vpd_stress <- function(vpd, a, b, m){
  out <- a * (b - m * (log(0.001) + log(vpd)))
  out <- ifelse(out>1, 1, out)
  out <- ifelse(out<0, 0, out)
  return(out)
}

data.frame(x = c(0, 2000)) %>% 
ggplot(aes(x = x)) +
        stat_function(fun = calc_vpd_stress, args = list(a=a, b=b, m=m)) +
  lims(y=c(0, 1)) +
  # geom_point(data = df, aes(x, y)) +
  labs(x = "VPD (Pa)", y = "Stress factor")
```

## Using rsofun

### Simulation settings

Create a site meta info table that contains all the site-specific information that is used to force site-simulations (e.g. starting year, number of simulations years, elevation, etc.). For FLUXNET2015 data, required meta info is provided by the `rsofun` package (data frame `rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015`).
```{r}
path_siteinfo <- "~/mct/siteinfo_mct_fluxnet2015.csv"
siteinfo <- rsofun::metainfo_Tier1_sites_kgclimate_fluxnet2015 %>% 
  dplyr::filter(!(sitename %in% c("DE-Akm", "IT-Ro1"))) %>%  # excluded because fapar data could not be downloaded (WEIRD)
  dplyr::filter(sitename != "FI-Sod") %>%  # excluded because some temperature data is missing
  dplyr::filter( c4 %in% c(FALSE, NA) & classid != "CRO" & classid != "WET" ) %>% 
  write_csv(path = path_siteinfo)
```

Now specify the simulation parameters that are identical for all site-scale simulations.
```{r}
params_siml <- list(
  spinup            = TRUE,
  spinupyears       = 10,
  recycle           = 1,
  soilmstress       = FALSE,
  tempstress        = FALSE,
  calc_aet_fapar_vpd = TRUE,
  in_ppfd           = TRUE,
  in_netrad         = FALSE,
  const_clim_year   = -9999,
  const_lu_year     = -9999,
  const_co2_year    = -9999,
  const_ndep_year   = -9999,
  const_nfert_year  = -9999,
  outdt             = 1,
  ltre              = FALSE,
  ltne              = FALSE,
  ltrd              = FALSE,
  ltnd              = FALSE,
  lgr3              = TRUE,
  lgn3              = FALSE,
  lgr4              = FALSE
  )
```

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters).
```{r}
settings_sims <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml)
```

### Define model parameters

First, let's do it by hand (calibration of parameters is shown later).
```{r}
params_modl <- list(
  kphio           = 0.04997714009213085,
  soilm_par_a     = 1.0,
  soilm_par_b     = 0.0,
  vpdstress_par_a = 0.1,
  vpdstress_par_b = 3.841307,
  vpdstress_par_m = 1.368163
  )
```

### Define soil parameters

For now, this is implemented as an illustration. Should be made site-specific. Is obsolete for now since soil moisture has no effect on ET, and since WHC is prescribed anyways.
```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

### Get input

First, define input settings.
```{r}
settings_input <-  list(
    data                     = NA,
    temperature              = "fluxnet2015",
    precipitation            = "fluxnet2015",
    vpd                      = "fluxnet2015",
    ppfd                     = "fluxnet2015",
    netrad                   = "fluxnet2015",  #  c("fluxnet2015", "watch_wfdei"),
    patm                     = "fluxnet2015",
    netrad                   = NA,
    cloudcover               = "cru",
    path_input               = "~/sofun_inputs/",
    path_watch_wfdei         = "~/data/watch_wfdei/",
    path_cru                 = "~/data/cru/ts_4.01/",
    path_MODIS_FPAR_MCD15A3H = "~/data/fluxnet_subsets/fapar_MODIS_FPAR_MCD15A3H_gee_MCD15A3H_fluxnet2015_gee_subset/",
    path_co2                 = "~/data/co2/cCO2_rcp85_const850-1765.csv",
    path_fluxnet2015         = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
    path_fluxnet2015_hh      = "~/data/FLUXNET-2015_Tier1/20191024/HH/",
    get_from_remote          = FALSE,
    settings_gee             = get_settings_gee( 
      bundle      = "fpar", 
      python_path = "/Users/benjaminstocker/Library/Enthought/Canopy_64bit/User/bin/python",
      gee_path    = "~/gee_subset/gee_subset/"
      ),
  fapar = "MODIS_FPAR_MCD15A3H",
  splined_fapar = TRUE
  )
```

Then, get the input data.
```{r message=FALSE, warning=FALSE, echo=FALSE}
ddf_input <- prepare_input_sofun(
  settings_input             = settings_input,
  settings_sims              = settings_sims,
  overwrite_csv_climate_lev1 = FALSE,
  overwrite_csv_climate_lev2 = FALSE,
  overwrite_csv_climate_lev3 = FALSE,
  overwrite_rdata_climate    = FALSE,
  overwrite_csv_fapar        = FALSE,
  verbose                    = FALSE
  )
```

### Run the model

Run the model for all the sites specified in the first step.
```{r}
df_drivers <- collect_drivers_sofun( 
  settings       = settings_sims, 
  forcing        = ddf_input, 
  df_soiltexture = df_soiltexture
  )

mod <- runread_sofun_f(
  df_drivers, 
  params_modl = params_modl, 
  makecheck = TRUE,
  parallel = FALSE
  )

mod$out_sofun[[1]] %>% 
  ggplot(aes(x=date, y=latenth)) +
  geom_line() + 
  labs(title = mod$sitename[[1]], subtitle = "SOFUN output")
```

### Calibrate

Define calibration settings.
```{r}
c4sites <- siteinfo %>% 
  dplyr::filter(c4) %>% 
  pull(sitename)

calibsites <- readr::read_csv( "~/data/flue/flue_stocker18nphyt.csv" ) %>%
              dplyr::select(site, cluster) %>% 
              dplyr::distinct() %>% 
              dplyr::filter( !is.na(cluster) ) %>%
              dplyr::filter( !(site %in% c4sites) ) %>%  # additionally exclude C4
              dplyr::filter(site %in% df_drivers$sitename) %>% 
              dplyr::pull(site)

settings_calib <- list(
  name             = "MCT",
  par              = list( vpdstress_par_a = list( lower=0.1,    upper=2,      init=1.0 ),
                           vpdstress_par_b = list( lower=0.1,    upper=1,      init=1 ),
                           vpdstress_par_m = list( lower=0.01,    upper=1,      init=0.5 ) ),
  method           = "gensa",
  targetvars       = c("latenth"),
  datasource       = list( latenth = "fluxnet2015" ),
  timescale        = list( latenth = "d" ),
  path_fluxnet2015 = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
  maxit            = 10,
  sitenames        = calibsites,
  metric           = "rmse",
  dir_results      = "~/mct/calib_results/"
)
```
 
Get calibration target data.
```{r warning=FALSE, messge=FALSE}
filn <- "./data/ddf_obs_calib.Rdata"
if (!file.exists(filn)){
  ddf_obs_calib <- get_obs_calib( 
    settings_calib = settings_calib, 
    dplyr::select(df_drivers, sitename, siteinfo) %>% tidyr::unnest(siteinfo), 
    settings_input
    )
  save(ddf_obs_calib, file = filn)
} else {
  load(filn)
}
```

First, as a test, calibrate using data from one site at a time and evaluate at that site, using the same data. To do this, we first need to define evaluation settings and get evaluation benchmarking data.
```{r}
mylist <- readr::read_csv("~/eval_pmodel/myselect_fluxnet2015.csv") %>% 
  dplyr::filter( use==1 ) %>% 
  dplyr::pull( Site )

settings_eval <- list(
  sitenames = settings_sims$sitename,
  sitenames_siteplots = mylist,
  agg = 8,
  path_fluxnet2015_d = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1d/original/unpacked/",
  path_fluxnet2015_w = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_7d/original/unpacked/",
  path_fluxnet2015_m = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1m/original/unpacked/",
  path_fluxnet2015_y = "~/data/FLUXNET-2015_Tier1/20160128/point-scale_none_1y/original/unpacked/",
  path_gepisat_d     = "~/data/gepisat/v3_fluxnet2015/daily_gpp/",
  benchmark = list( latenth = c("fluxnet2015") ),
  remove_premodis = TRUE
  )
```

Get evaluation data (benchmarking data).
```{r}
filn <- "./obs_eval.Rdata"
if (file.exists(filn)){
  load(filn)
} else {
  obs_eval  <- get_obs_eval( 
    settings_eval = settings_eval, 
    settings_sims = settings_sims, 
    overwrite     = TRUE, 
    light         = TRUE,
    add_forcing   = FALSE
  )
  save(obs_eval, file = filn)
} 
```

#### Calibrate/evaluate by site

Site-by-site calibration/evaluation (not out-of-bag!)
```{r}
filn <- "./data/out_sitebysite.Rdata"
if (file.exists(filn)){
  load(filn)
} else {
  out_sitebysite <- sbs_calib_eval_sofun( 
    settings_calib, 
    settings_eval, 
    settings_sims, 
    df_drivers = df_drivers,
    ddf_obs_calib = ddf_obs_calib, 
    obs_eval = obs_eval,
    overwrite = TRUE
    )
  save(out_sitebysite, file = filn)
}
```

Extract site-level parameter values and metrics and plot.
```{r}

## extract metrics from individual left-out calibrations/evaluations
source("~/eval_pmodel/extract_metric_oob.R")
df_metrics <- extract_metric_oob(out_sitebysite, metric = "rsq", benchmarkvar = "latenth") %>% 
  left_join(extract_metric_oob(out_sitebysite, metric = "rmse", benchmarkvar = "latenth"), by = "sitename")

# get parameters from individual left-out calibrations
source("~/eval_pmodel/extract_params_oob.R")
df_params <- extract_params_oob(param = "vpdstress_par_a", path = "~/mct/calib_results/sitebysite_MCT", sitenames = df_metrics$sitename)

# a parameter
gg_a <- df_params %>% 
  ggplot() +
  geom_histogram(aes(x=vpdstress_par_a, y = ..count..),
    color = "black", alpha = 0.3, 
    position="identity") +
  geom_vline(aes(xintercept = mean(df_params$kphio)), col = "red", linetype = "dashed") +  # Mean across out-of-bag evaluations
  theme_classic() +
  labs(x = latex2exp::TeX("$\\widehat{a}$"), y = "Count")  # bquote(varphi[0])

## b-parameter
gg_b <- df_params %>% 
  ggplot() +
  geom_histogram(aes(x=vpdstress_par_b, y = ..count..),
    color = "black", alpha = 0.3, 
    position="identity") +
  geom_vline(aes(xintercept = mean(df_params$kphio)), col = "red", linetype = "dashed") +  # Mean across out-of-bag evaluations
  theme_classic() +
  labs(x = latex2exp::TeX("$\\widehat{b}$"), y = "Count")  # bquote(varphi[0])

## m-parameter
gg_m <- df_params %>% 
  ggplot() +
  geom_histogram(aes(x=vpdstress_par_m, y = ..count..),
    color = "black", alpha = 0.3, 
    position="identity") +
  geom_vline(aes(xintercept = mean(df_params$kphio)), col = "red", linetype = "dashed") +  # Mean across out-of-bag evaluations
  theme_classic() +
  labs(x = latex2exp::TeX("$\\widehat{m}$"), y = "Count")  # bquote(varphi[0])

## R2
gg_rsq <- df_metrics %>% 
  ggplot() +
  geom_histogram(aes(x = rsq, y = ..count..),
    color = "black", alpha = 0.3, binwidth = 0.05, 
    position="identity") +
  # geom_vline(aes(xintercept = out_eval_FULL$latenth$fluxnet2015$metrics$xdaily_pooled$rsq), col = "red") +  # All sites simultaneously
  geom_vline(aes(xintercept = mean(df_metrics$rsq)), col = "red", linetype = "dashed") +  # Mean across out-of-bag evaluations
  theme_classic() +
  labs(x = bquote(italic(R)^2), y = "Count")
  # scale_linetype_manual(name = "", values = c(all = "solid", mean = "dashed"), labels = c(all = "All sites simultaneously", mean = "Mean across out-of-bag evaluations")) +
  # theme(legend.position="top")

## RMSE
gg_rmse <- df_metrics %>% 
  ggplot() +
  geom_histogram(aes(x = rmse, y = ..count..),
    color = "black", alpha = 0.3, 
    position="identity") +
  # geom_vline(aes(xintercept = out_eval_FULL$gpp$fluxnet2015$metrics$xdaily_pooled$rmse, linetype = "all"), col = "red") +
  geom_vline(aes(xintercept = mean(df_metrics$rmse), linetype = "mean"), col = "red") + 
  theme_classic() +
  labs(x = expression( paste("RMSE (g C m"^-2, " d"^-1, ")" ) ), y = "Count") +
  scale_linetype_manual(name = "", values = c(all = "solid", mean = "dashed"), labels = c(all = "All sites simultaneously", mean = "Mean across out-of-bag evaluations"))
  # theme(legend.position = c(0.6,0.85))

toprow <- cowplot::plot_grid(gg_a, gg_b, gg_m, labels = "auto", ncol = 3)
bottomrow <- cowplot::plot_grid(gg_rsq, gg_rmse, labels = c("d", "e"), ncol = 2, rel_widths = c(1,2))

cowplot::plot_grid(toprow, bottomrow, nrow = 2)
```

Plot the VPD-stress function for the different sites.
```{r}
plot_vpdstress_bysite <- function(a, b, m, sitename){
  df <- tibble(
    x = seq(0, 2000, by = 20)
    ) %>% 
    mutate(y = calc_vpd_stress(x, a, b, m) )
  
  gg <- data.frame(x = c(0, 2000)) %>% 
    ggplot(aes(x = x)) +
          stat_function(fun = calc_vpd_stress, args = list(a=a, b=b, m=m)) +
    lims(y=c(0, 1)) +
    geom_point(data = df, aes(x, y)) +
    labs(x = "VPD (Pa)", y = "Stress factor", title = sitename)
  return(gg)
}
out_vpdstress <- df_params %>% 
  rename(a = vpdstress_par_a, b = vpdstress_par_b, m = vpdstress_par_m) %>% 
  mutate(out_gg = purrr::pmap(., plot_vpdstress_bysite))
out_vpdstress$out_gg
```

Some seasonal plots.
```{r}
plot_meandoy <- function(sitename, out_sitebysite){
  gg <- out_sitebysite[[sitename]]$latenth$fluxnet2015$data$meandoydf %>% 
    ggplot() +
    geom_ribbon(
      aes(x = doy, ymin = obs_min, ymax = obs_max), fill = "black", alpha = 0.3
      ) +
    geom_line(aes(x = doy, y = mod_mean), size = 0.4) +
    theme_gray() +
    labs(title = sitename)
  return(gg)
}
gg_list <- purrr::map(
  as.list(names(out_sitebysite)),
  plot_meandoy(., out_sitebysite)
  )
gg_list[[1]]
gg_list[[2]]
gg_list[[3]]
gg_list[[4]]
gg_list[[5]]
```

#### Calibrate with pooled data

```{r}
set.seed(1982)
settings_calib <- calib_sofun( 
  settings_calib, 
  df_drivers, 
  ddf_obs = ddf_obs_calib 
  )
```

The calibrated parameters are returned by `calib_sofun()` as part of the list:
```{r}
print(settings_calib$par_opt)
```

This gives a stress function that looks like this:
```{r}
df <- tibble(
  x = seq(0, 2000, by = 20)
  ) %>% 
  mutate(y = calc_vpd_stress(x, a, b, m) )

data.frame(x = c(0, 2000)) %>% 
  ggplot(aes(x = x)) +
        stat_function(fun = calc_vpd_stress, args = list(a=settings_calib$par_opt["vpdstress_par_a"], b=settings_calib$par_opt["vpdstress_par_b"], m=settings_calib$par_opt["vpdstress_par_m"])) +
  lims(y=c(0, 1)) +
  labs(x = "VPD (Pa)", y = "Stress factor", title = "ALL")
```

### Evaluate

Run the model once again with these parameters and evaluate results.

Now run the model with calibrated parameters.
```{r}
params_modl <- list(
  kphio           = 0.04997714009213085,
  soilm_par_a     = 1.0,
  soilm_par_b     = 0.0,
  vpdstress_par_a = settings_calib$par_opt[["vpdstress_par_a"]],
  vpdstress_par_b = settings_calib$par_opt[["vpdstress_par_b"]],
  vpdstress_par_m = settings_calib$par_opt[["vpdstress_par_m"]]
  )

mod <- runread_sofun_f(
  df_drivers, 
  params_modl = params_modl, 
  makecheck = TRUE
  ) %>% 
  rename(id = sitename) %>% 
  mutate(ntsteps = purrr::map_int(out_sofun, ~nrow(.))) %>% 
  filter(ntsteps>1) %>% 
  unnest(out_sofun)
```

Get evaluation
```{r}
out_eval <- eval_sofun( 
  mod, 
  settings_eval, 
  settings_sims, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

Modelled vs. Observed
```{r}
out_modobs <- out_eval$latenth$fluxnet2015$data$xdf %>% rbeni::analyse_modobs2("mod", "obs", type = "heat")
out_modobs$gg
```

Some seasonal plots.
```{r}
plot_meandoy <- function(mysitename, out_eval){
  gg <- out_eval$latenth$fluxnet2015$data$meandoydf %>% 
    dplyr::filter(sitename==mysitename) %>% 
    ggplot() +
    geom_ribbon(
      aes(x = doy, ymin = obs_min, ymax = obs_max), fill = "black", alpha = 0.3
      ) +
    geom_line(aes(x = doy, y = mod_mean), size = 0.4) +
    theme_gray() +
    labs(title = mysitename)
  return(gg)
}

gg_list <- purrr::map(
  as.list(out_eval$latenth$fluxnet2015$data$meandoydf %>% pull(sitename) %>% unique()),
  ~plot_meandoy(., out_eval)
  )

gg_list[[1]]
gg_list[[2]]
gg_list[[3]]
gg_list[[4]]
gg_list[[5]]
```

