---
title: "Create Supplementary Information"
author: "Beni Stocker"
date: "2/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rbeni)
```

## Testing $S_0$ diagnosing

As testing sites, use sites from the rooting depth dataset (RSIP).

```{r cars}
df_sites <- read_csv("data/df_sites_rsip.csv")
```

Extract information of S0 diagnostic for given locations.

Sample 30 sites out of the 1705.
```{r}
set.seed(1982)
use_biomes <- df_sites |> 
  group_by(biome) |> 
  summarise(n = n()) |> 
  dplyr::filter(n >=5 & !is.na(biome)) |> 
  pull(biome)

df_sites_sampled <- df_sites |> 
  dplyr::filter(biome %in% use_biomes) |> 
  group_by(biome) |> 
  group_split() |> 
  purrr::map(~sample_n(., 3)) |> 
  bind_rows()
```

Plot sites on map
```{r}
library(jcolors)
nbiomes <- df_sites_sampled |> 
  drop_na(biome_name) |>
  pull(biome_name) |> 
  unique() |> 
  length()

plot_map_simpl() +
  geom_point(data = df_sites_sampled,
             aes(x = lon, y = lat, color = biome_name)) +
  # scale_color_jcolors(palette = "pal8")
  # scale_color_brewer(palette = "Paired")
  scale_colour_manual(values = myjcolors(nbiomes), name = "Biome name")
```

Determine all chunks (latitudinal bands)
```{r}
get_ilon <- function(lon){
  lon_hires <- seq(-179.975, 179.975, by = 0.05)
  ilon <- which.min(abs(lon - lon_hires))
  return(ilon)  
}

df_sites_sampled <- df_sites_sampled |> 
  mutate(ilon = purrr::map_dbl(lon, ~get_ilon(.)))
```

Get all data for chunks.
```{r}
source("R/calc_cwd_et0_byilon.R")
system("mkdir -p ~/mct/data/df_cwd_et0_3")
filn <- "data/df_test_cwd_et0.rds"

if (!file.exists(filn)){
  df <- df_sites_sampled |> 
    slice(1) |> 
    mutate(out = purrr::map2(ilon, 
                             lat, 
                             ~calc_cwd_et0_byilon(.x, 
                                                  drop_data = FALSE, 
                                                  dirn = "~/mct/data/df_cwd_et0_3/", 
                                                  verbose = FALSE, 
                                                  overwrite = TRUE, 
                                                  siteinfo = NULL, 
                                                  do_plot = TRUE, 
                                                  use_lat = .y
                                                  )))
  
  saveRDS(df, file = filn)
} else {
  df <- readRDS(filn)
}
```

Visualise
```{r}
# myplot <- function(n, df){
#   print(n)  
#   df$out[[n]]$gg_fet
# }
# purrr::map(as.list(seq(36)), ~myplot(., df))

## example for no decline
gg1 <- df$out[[2]]$gg_fet[[1]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()

## example for S0 diagnosed (9, 34, 21)
gg2 <- df$out[[34]]$gg_fet[[1]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()

gg3 <- df$out[[21]]$gg_fet[[1]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()

## example for flattening
gg4 <- df$out[[31]]$gg_fet[[1]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()

cowplot::plot_grid(gg1, gg4, gg2, gg3, nrow = 2, labels = c('a', 'b', 'c', 'd'))
ggsave("fig/plot_test_s0_diag.pdf", width = 12, height = 8)
ggsave("fig/plot_test_s0_diag.png", width = 12, height = 8)
```

Visualise
```{r}
tmp <- df |> 
  dplyr::select(-lon, -lat) |> 
  unnest(out) |> 
  mutate(gg = purrr::map(out, ~pull(slice(., 1), gg_fet)),
         cwd_lue0_fet = purrr::map_dbl(out, ~pull(., cwd_lue0_fet)))
  
## example for no decline
gg1 <- tmp$gg_fet[[2]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()

## example for S0 diagnosed (9, 34, 21)
gg2 <- tmp$gg_fet[[32]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()

gg3 <- tmp$gg_fet[[21]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()  
  
## example for flattening
gg4 <- tmp$gg_fet[[29]] +
  labs(title = NULL, y = "EF (unitless)") +
  theme_classic()  

cowplot::plot_grid(gg1, gg4, gg2, gg3, nrow = 2, labels = c('a', 'b', 'c', 'd'))

## retain selected
tmp2 <- tmp |> 
  slice(c(2, 32, 21, 29))

# ## s_def diagnostic
# df$out[[1]]$gg_fet
# 
# ## water balance time series
# df$out[[1]]$data[[1]] |> 
#   ggplot(aes(time, bal)) + 
#   geom_line()
# 
# df$out[[1]]$data[[1]] |> 
#   ggplot(aes(deficit, fet)) + 
#   geom_point()
# 
# df$out[[1]]$data[[1]] |> 
#   ggplot(aes(NR, et)) + 
#   geom_point() +
#   geom_smooth(method = "lm")
```

Plot the same from the model outputs.
```{r}
# ## determine chunk used for the four sites in tmp2
# df_sites_ichunk <- read_csv("data/df_sites_rsip.csv") |>
#   mutate(idx = 1:n()) |>
#   mutate(chunk = rep(1:as.integer(50),
#                      each = (nrow(.)/as.integer(50)), len = nrow(.)))
# 
# tmp2 <- tmp2 |> 
#   left_join(
#     df_sites_ichunk |> 
#       select(lon, lat, Dr, chunk)
#   )

read_myfile <- function(ichunk, use_whc){
  path <- "./data/out_rsofun_cwdx/"
  filename <- file.path(path, paste0("out_rsofun_cwdx_whc_", as.character(use_whc), "_ichunk_", as.character(ichunk), ".rds"))
  if (file.exists(filename)){
    df <- readRDS(filename)
    return(df)
  } else {
    return(tibble())
  }
}

set.seed(1982)
df_rsofun <- purrr::map_dfr(
      as.list(sample(seq(50), 3)),
      ~read_myfile(., "200")
      ) |> 
      mutate(setup = "whc_200")

# ## example of S0 detected
# df_rsofun$out_cwd_lue0[[20]]$gg +
#   geom_vline(xintercept = df_rsofun$whc[20])

```

```{r}
purrr::map(seq(1:30), ~{print(df_rsofun$out_cwd_lue0[[.]]$gg)})

```

## Magnitudes of EF

Read FLUXNET2015 data
```{r}
library(ingestr)
ddf_fluxnet <- ingest(
  siteinfo  = siteinfo_fluxnet2015,
  source    = "fluxnet",
  getvars   = list(netrad = "NETRAD", et = "LE_F_MDS"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",  # adjust this with your local path
  settings  = list(
    getswc       = FALSE,
    filter_ntdt  = FALSE,
    remove_neg   = FALSE,
    threshold_LE = 0.8
    ),
  timescale = "d",
  verbose = TRUE
  )

## take 95% quantile of EF by site
tmp <- ddf_fluxnet |> 
  mutate(data = purrr::map(data, ~mutate(., ef = remove_outliers(et/netrad)))) |> 
  unnest(data) |> 
  group_by(sitename) |> 
  summarise(ef_q95 = quantile(ef, probs = 0.90, na.rm = TRUE))

tmp |> 
  ggplot(aes(ef_q95, ..count..)) +
  geom_histogram() + 
  theme_classic() +
  labs(x = "Maximum evaporative fraction (unitless)", y = "Count") +
  xlim(0, 2) +
  geom_vline(xintercept = 1.0, linetype = "dotted")
ggsave("fig/max_ef_fluxnet2015.pdf", width = 6, height = 4)
ggsave("fig/max_ef_fluxnet2015.png", width = 6, height = 4)
```


## CWD demo

```{r}
source("R/test_cwd_tseries.R")
# load("~/mct/data/df_cwdx/df_cwdx_ilon_5000.RData")

idxc <- 1
idxa <- 500
idxb <- 300 # 100

gg1 <- test_cwd_tseries(
  df$out_mct[[idxa]]$mct, 
  sitename = NA, 
  filter_years = NA, 
  title = paste0(as.character(df$lon[idxa]), "°E, ", as.character(df$lat[idxa]), "°N")
  )

gg2 <- test_cwd_tseries(
  df$out_mct[[idxb]]$mct, 
  sitename = NA, 
  filter_years = NA, 
  title = paste0(as.character(df$lon[idxb]), "°E, ", as.character(df$lat[idxb]), "°N")
  )

gg3 <- test_cwd_tseries(
  df$out_mct[[idxc]]$mct, 
  sitename = NA, 
  filter_years = NA, 
  title = paste0(as.character(df$lon[idxc]), "°E, ", as.character(df$lat[idxc]), "°N")
  )

df_sites <- tibble(
  sitename = c("a", "b", "c"),
  lon = c(df$lon[idxa], df$lon[idxb], df$lon[idxc]),
  lat = c(df$lat[idxa], df$lat[idxb], df$lat[idxc])
)

library(rbeni)
library(ggrepel)

gg0 <- plot_map_simpl() +
  geom_text(data = df_sites, aes(lon, lat, label = sitename), color = "red")

cowplot::plot_grid(gg0, gg1, gg2, gg3, labels = c("", "a", "b", "c"), nrow = 2)
ggsave("fig/cwd_tseries.pdf", width = 12, height = 7)
ggsave("fig/cwd_tseries.png", width = 12, height = 7)
```

## S0 versus CTI 

How does S0 vary along topographic gradients? Select limited region (not too different climate) and plot S0 distribution (boxplot) along CTI bins.
```{r}
load("data/df_cwdx_10_20_40.RData") # loads 'df', created by rscript_collect_cwdx.R

df_cwdx <- df |>
  mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3))
```

### Add CTI data

```{r}
## combine S_cwdxXX and GTI into one big data frame
df_cwdx <- df |> 
  mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)) |> 
  left_join(
    nc_to_df("~/data/gti_marthews/ga2_0_05deg_median.nc", varnam = "gti") |>
      drop_na() |>
      mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)) |> 
      rename(gti_q50 = gti),
    by = c("lon", "lat")
  ) |> 
  left_join(
    nc_to_df("~/data/gti_marthews/ga2_0_05deg_q10.nc", varnam = "gti") |>
      drop_na() |>
      mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)) |> 
      rename(gti_q10 = gti),
    by = c("lon", "lat")
  ) |> 
  left_join(
    nc_to_df("~/data/gti_marthews/ga2_0_05deg_q90.nc", varnam = "gti") |>
      drop_na() |>
      mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)) |> 
      rename(gti_q90 = gti),
    by = c("lon", "lat")
  ) |> 
  mutate(gti_q10 = ifelse(gti_q10 < 0, NA, gti_q10),
         gti_q50 = ifelse(gti_q50 < 0, NA, gti_q50),
         gti_q90 = ifelse(gti_q90 < 0, NA, gti_q90))
```

### Add elevation data

```{r}
## combine S_cwdxXX and GTI into one big data frame
df_cwdx <- df_cwdx %>% 
  left_join(
    nc_to_df("~/data/etopo/ETOPO1_Bed_g_mean_0_05deg.nc", varnam = "elv") %>%
      drop_na() %>%
      mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)),
    by = c("lon", "lat")) %>% 
  mutate(elv = ifelse(elv < 0, NA, elv))
```

### Apply vegetation mask

```{r}
load("data/df_vegmask.RData")  # loads df_vegmask

df_cwdx <- df_cwdx |>
  left_join(df_vegmask,
            by = c("lon", "lat")) |>
  mutate(cwdx80 = ifelse(nonveg > 99, NA, cwdx80)) |> 
  mutate(cwdx80 = ifelse(water > 99, NA, cwdx80)) |> 
  mutate(cwdx80 = ifelse(snowandice > 99, NA, cwdx80)) 
```
    
S0 in CTI and elevation bins
```{r}
gti_breaks <- 0:15
elv_breaks <- seq(0, 9, by = 0.5)

df_cwdx <- df_cwdx %>%
  mutate(elv = elv/1000) %>% 
  ungroup() %>%
  mutate(igti_q10 = cut(gti_q10,
                    breaks = gti_breaks),
         igti_q50 = cut(gti_q50,
                    breaks = gti_breaks),
         igti_q90 = cut(gti_q90,
                    breaks = gti_breaks)) %>%
  mutate(ielv = cut(elv,
                    breaks = elv_breaks))
```  

### Central asia
    
```{r}
df_box1 <- tibble(
  long = c(50, 65, 65, 50, 50), lat = c(26, 26, 35, 35, 26),
  order = 1:5, group = rep(1, 5)
  )

df_box2 <- tibble(
  long = c(76, 84, 84, 76, 76), lat = c(26, 26, 31, 31, 26),
  order = 1:5, group = rep(1, 5)
  )

## central asia
gg_cwdx <- df_cwdx |> 
  drop_na(cwdx80) |> 
  plot_map4(varnam = "cwdx80", lonmin = 45, lonmax = 95, latmin = 25, latmax = 47.5, 
            breaks = c(seq(0, 100, by = 20), 150, 200, 300, 500, 700, 900, 1200, Inf), 
            spacing = "constant",
            combine = FALSE, 
            colorscale = "batlowK", 
            legend_title = "(mm)",
            expand_size_y = 0.5,
            hillshade = TRUE, rivers = TRUE, lakes = TRUE,
            scale = 50
            )

## add red box for zoom
gg_cwdx$ggmap <- gg_cwdx$ggmap + 
  labs(title = expression(paste(italic("S")[CWDX80]))) +
  geom_path(aes(x = long, y = lat, group = group), data = df_box1, size = 0.5, color = "red") +
  geom_path(aes(x = long, y = lat, group = group), data = df_box2, size = 0.5, color = "royalblue")

cowplot::plot_grid(gg_cwdx$ggmap, gg_cwdx$gglegend, ncol = 2, rel_widths = c(1, 0.2))
```

Map CTI
```{r}
## central asia
gg_cti <- df_cwdx |> 
  drop_na(gti_q90) |> 
  plot_map4(varnam = "gti_q90", lonmin = 45, lonmax = 95, latmin = 25, latmax = 47.5, 
            breaks = 0:15, 
            spacing = "constant",
            combine = FALSE, 
            colorscale = "batlowK", legend_title = "(unitless)",
            expand_size_y = 0.5,
            hillshade = TRUE, rivers = TRUE, lakes = TRUE,
            scale = 50
            )
gg_cti$ggmap <- gg_cti$ggmap + 
  labs(title = "CTI, 90% quantile") + 
  geom_path(aes(x = long, y = lat, group = group), data = df_box1, size = 0.5, color = "red") +
  geom_path(aes(x = long, y = lat, group = group), data = df_box2, size = 0.5, color = "royalblue")
```

Map Elevation
```{r}
## central asia
gg_elv <- df_cwdx %>% 
  plot_map4(varnam = "elv", lonmin = 45, lonmax = 95, latmin = 25, latmax = 47.5, 
            breaks = seq(0, 8.500, by = 0.500), 
            spacing = "constant",
            combine = FALSE, 
            colorscale = "batlow", invert = 1, legend_title = "(km)",
            expand_size_y = 0.5,
            hillshade = TRUE, rivers = TRUE, lakes = TRUE,
            scale = "medium"
            )
gg_elv$ggmap <- gg_elv$ggmap + 
  labs(title = "Elevation") + 
  geom_path(aes(x = long, y = lat, group = group), data = df_box1, size = 0.5, color = "red") +
  geom_path(aes(x = long, y = lat, group = group), data = df_box2, size = 0.5, color = "royalblue")

gg_elv$ggmap
```

Red box: `long = c(50, 65, 65, 50, 50), lat = c(26, 26, 35, 35, 26)`
```{r}
n_fun <- function(x){
  return(data.frame(y = 750,
                    label = length(x)))
}

gg_boxplot <- df_cwdx |> 
  filter(lon > 50 & lon < 65 & lat > 26 & lat < 35 & !is.na(igti_q90)) |>
  ggplot(aes(igti_q90, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  stat_summary(fun.data = n_fun, geom = "text", hjust = 0.5, size = 2.5) +
  ylim(0, 750) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "CTI bin", title = "CTI", subtitle = "Red box")

gg_boxplot1_cwdx_casia

gg_boxplot1_elv_casia <- df_cwdx %>% 
  filter(lon > 50 & lon < 65 & lat > 26 & lat < 35 & !is.na(ielv)) %>%
  ggplot(aes(ielv, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  ylim(0, 600) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "Elevation bin (km)", title = "Elevation", subtitle = "Red box")

gg_boxplot1_elv_casia
```

Blue box: `long = c(76, 84, 84, 76, 76), lat = c(26, 26, 31, 31, 26)`
```{r}
gg_boxplot2_cwdx_casia <- df_cwdx %>% 
  filter(lon > 76 & lon < 84 & lat > 26 & lat < 31 & !is.na(igti_q90)) %>%
  ggplot(aes(igti_q90, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  stat_summary(fun.data = n_fun, geom = "text", hjust = 0.5, size = 2.5) +
  ylim(0, 750) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "CTI bin", title = "CTI", subtitle = "Blue box")

gg_boxplot2_cwdx_casia

gg_boxplot2_elv_casia <- df_cwdx %>% 
  filter(lon > 76 & lon < 84 & lat > 26 & lat < 31 & !is.na(ielv)) %>%
  ggplot(aes(ielv, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  ylim(0, 600) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "Elevation bin (km)", title = "Elevation", subtitle = "Blue box")

gg_boxplot2_elv_casia
```

Publication figure
```{r}
top <- cowplot::plot_grid(gg_cwdx$ggmap, gg_cwdx$gglegend, 
                          gg_cti$ggmap, gg_cti$gglegend,
                          gg_elv$ggmap, gg_elv$gglegend,
                          ncol = 2, rel_widths = c(1, 0.12), labels = c('a', '', 'b', '', 'c', ''))
btm <- cowplot::plot_grid(gg_boxplot1_cwdx_casia, gg_boxplot2_cwdx_casia,
                          gg_boxplot1_elv_casia, gg_boxplot2_elv_casia,
                          ncol = 2, labels = c('d', 'e', 'f', 'g'))
cowplot::plot_grid(top, btm, ncol = 1, rel_heights = c(3, 0.85))
ggsave("fig/fig_topography_casia.png", width = 10, height = 19)
ggsave("fig/fig_topography_casia.pdf", width = 10, height = 19)
```

### Western USA
    
```{r}
df_box1 <- tibble(
  long = c(-113, -119, -119, -113, -113), lat = c(38, 38, 42, 42, 38),
  order = 1:5, group = rep(1, 5)
  )

df_box2 <- tibble(
  long = c(-118, -120.5, -120.5, -118, -118), lat = c(35, 35, 37, 37, 35),
  order = 1:5, group = rep(1, 5)
  )

## W USA
gg_cwdx <- df_cwdx |> 
  drop_na(cwdx80) |> 
  plot_map4(varnam = "cwdx80", 
            lonmin = -125, lonmax = -95, latmin = 32.5, latmax = 47,
            breaks = c(seq(0, 100, by = 20), 150, 200, 300, 500, 700, 900, 1200, Inf), 
            spacing = "constant",
            combine = FALSE, 
            colorscale = "batlowK", legend_title = "(mm)",
            expand_size_y = 0.5,
            hillshade = TRUE, rivers = TRUE, lakes = TRUE,
            scale = 50
            )

## add red box for zoom
gg_cwdx$ggmap <- gg_cwdx$ggmap + 
  labs(title = expression(paste(italic("S")[CWDX80]))) +
  geom_path(aes(x = long, y = lat, group = group), data = df_box1, size = 0.5, color = "red") +
  geom_path(aes(x = long, y = lat, group = group), data = df_box2, size = 0.5, color = "royalblue")

# cowplot::plot_grid(gg_cwdx$ggmap, gg_cwdx$gglegend, ncol = 2, rel_widths = c(1, 0.2))
```

Plot CTI
```{r}
## central asia
gg_cti <- df_cwdx |> 
  drop_na(gti_q90) |> 
  plot_map4(varnam = "gti_q90", 
            lonmin = -125, lonmax = -95, latmin = 32.5, latmax = 47,
            breaks = 0:15, 
            spacing = "constant",
            combine = FALSE, 
            colorscale = "batlowK", legend_title = "(unitless)",
            expand_size_y = 0.5,
            hillshade = TRUE, rivers = TRUE, lakes = TRUE,
            scale = 50
            )
gg_cti$ggmap <- gg_cti$ggmap + 
  labs(title = "CTI, 90% quantile") + 
  geom_path(aes(x = long, y = lat, group = group), data = df_box1, size = 0.5, color = "red") +
  geom_path(aes(x = long, y = lat, group = group), data = df_box2, size = 0.5, color = "royalblue")
```

Map Elevation
```{r}
## central asia
gg_elv <- df_cwdx %>% 
  plot_map4(varnam = "elv", 
            lonmin = -125, lonmax = -95, latmin = 32.5, latmax = 47,
            breaks = seq(0, 4.000, by = 0.200), 
            spacing = "constant",
            combine = FALSE, 
            colorscale = "batlow", invert = 1, legend_title = "(km)",
            expand_size_y = 0.5,
            hillshade = TRUE, rivers = TRUE, lakes = TRUE,
            scale = "medium"
            )
gg_elv$ggmap <- gg_elv$ggmap + 
  labs(title = "Elevation") + 
  geom_path(aes(x = long, y = lat, group = group), data = df_box1, size = 0.5, color = "red") +
  geom_path(aes(x = long, y = lat, group = group), data = df_box2, size = 0.5, color = "royalblue")

gg_elv$ggmap
```

Red box 
```{r}
n_fun <- function(x){
  return(data.frame(y = 700,
                    label = length(x)))
}

gg_boxplot1_cwdx_wusa <- df_cwdx %>% 
  filter(lon > -119 & lon < -113 & lat > 38 & lat < 42 & !is.na(igti_q90)) %>%
  ggplot(aes(igti_q90, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  stat_summary(fun.data = n_fun, geom = "text", hjust = 0.5, size = 2.5) +
  ylim(0, 700) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "CTI bin", title = "CTI", subtitle = "Red box")

gg_boxplot1_cwdx_wusa

n_fun <- function(x){
  return(data.frame(y = 600,
                    label = length(x)))
}

gg_boxplot1_elv_wusa <- df_cwdx %>% 
  filter(lon > -119 & lon < -113 & lat > 38 & lat < 42 & !is.na(ielv)) %>%
  ggplot(aes(ielv, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  stat_summary(fun.data = n_fun, geom = "text", hjust = 0.5, size = 2.5) +
  ylim(0, 600) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "Elevation bin (km)", title = "Elevation", subtitle = "Red box")

gg_boxplot1_elv_wusa
```

Blue box `long = c(-118, -120.5, -120.5, -118, -118), lat = c(35, 35, 37, 37, 35),`
```{r}
n_fun <- function(x){
  return(data.frame(y = 700,
                    label = length(x)))
}
gg_boxplot2_cwdx_wusa <- df_cwdx %>% 
  filter(lon > -120.5 & lon < -118 & lat > 35 & lat < 37 & !is.na(igti_q90)) %>%
  ggplot(aes(igti_q90, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  stat_summary(fun.data = n_fun, geom = "text", hjust = 0.5, size = 2.5) +
  ylim(0, 700) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "CTI bin", title = "CTI", subtitle = "Blue box")

gg_boxplot2_cwdx_wusa

n_fun <- function(x){
  return(data.frame(y = 600,
                    label = length(x)))
}
gg_boxplot2_elv_wusa <- df_cwdx %>% 
  filter(lon > -120.5 & lon < -118 & lat > 35 & lat < 37 & !is.na(ielv)) %>%
  ggplot(aes(ielv, cwdx80)) +
  geom_boxplot(outlier.shape = NA, fill = "grey70") +
  stat_summary(fun.data = n_fun, geom = "text", hjust = 0.5, size = 2.5) +
  ylim(0, 600) +
  theme_classic() +
  labs(y = expression(paste(italic("S")[CWDX80])), x = "Elevation bin (km)", title = "Elevation", subtitle = "Blue box")

gg_boxplot2_elv_wusa
```


Publication figure
```{r}
top <- cowplot::plot_grid(gg_cwdx$ggmap, gg_cwdx$gglegend, 
                          gg_cti$ggmap, gg_cti$gglegend,
                          gg_elv$ggmap, gg_elv$gglegend,
                          ncol = 2, rel_widths = c(1, 0.12), labels = c('a', '', 'b', '', 'c', ''))
btm <- cowplot::plot_grid(gg_boxplot1_cwdx_wusa, gg_boxplot2_cwdx_wusa,
                          gg_boxplot1_elv_wusa, gg_boxplot2_elv_wusa,
                          ncol = 2, labels = c('d', 'e', 'f', 'g'))
cowplot::plot_grid(top, btm, ncol = 1, rel_heights = c(3, 0.85))
ggsave("fig/fig_topography_wusa.png", width = 10, height = 19)
ggsave("fig/fig_topography_wusa.pdf", width = 10, height = 19)
```


<!-- Publication figure -->
<!-- ```{r} -->
<!-- cowplot::plot_grid(gg_cwdx$ggmap, gg_cwdx$gglegend, gg_cti$ggmap, gg_cti$gglegend, gg_boxplot, -->
<!--                    ncol = 2, rel_widths = c(1, 0.12), rel_heights = c(1, 1, 0.4), labels = c('a', '', 'b', '', 'c', '')) -->
<!-- ggsave("fig/fig_topography_wusa.png", width = 10, height = 12) -->
<!-- ``` -->
