---
title: "Evaluation with ET data"
author: "Beni Stocker"
date: "5/17/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

See `~/rsofun/vignettes/splash.Rmd` for how site-scale simulations were done.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(ggplot2)
library(lubridate)
library(rbeni)
df_flue <- read_csv("~/data/fLUE/flue_stocker18nphyt.csv")
```

## "Correcting fAPAR data"

fAPAR really doesn't go to zero in drought-affected regions. This has caused an overestimation of GPP in respective sites as identified in an earlier analysis. Therefore, we correct fAPAR data
```{r  message=FALSE, warning=FALSE}
## run vignette 'splash_fluxnet2015.Rmd' in rsofun repo
load("~/mct/data/obs_eval_v2.RData")
df_flue_dry <- df_flue %>% filter(cluster %in% c("cGR", "cDD"))
ddf_obs_dry <- obs_eval$ddf %>% 
  filter(sitename %in% df_flue_dry$site)

fpar_absmin <- 0.12
ddf_obs_dry %>% 
  ggplot(aes(fapar, stat(density))) +
  geom_histogram( ) +
  geom_density() +
  geom_vline(xintercept = fpar_absmin, color="red") +
  xlim(0,1)
```

Do the same thing with EVI.
```{r}
load("~/mct/data/obs_eval_v2_EVI.RData")
obs_eval_evi <- obs_eval

ddf_obs_dry_evi <- obs_eval_evi$ddf %>% 
  filter(sitename %in% df_flue_dry$site)

evi_absmin <- 0.1
ddf_obs_dry_evi %>% 
  ggplot(aes(fapar, stat(density))) +
  geom_histogram( ) +
  geom_density() +
  geom_vline(xintercept = evi_absmin, color="red") +
  xlim(0,1)
```


Looking at mean fAPAR seasonality by sites it seems like the "absolute minimum" is around 0.15 and not 0.
```{r  message=FALSE, warning=FALSE}
df_meandoy <- ddf_obs_dry %>% 
  mutate(doy = yday(date)) %>% 
  group_by(sitename, doy) %>% 
  summarise( fapar_mean = mean( fapar, na.rm=TRUE ), 
             fapar_min  = min( fapar, na.rm=TRUE ), 
             fapar_max  = max( fapar, na.rm=TRUE )
             ) %>%
  mutate( fapar_min = ifelse( is.infinite(fapar_min), NA, fapar_min ), 
          fapar_max = ifelse( is.infinite(fapar_max), NA, fapar_max )) %>% 
  left_join(
    dplyr::select(ddf_obs_dry_evi, sitename, date, evi = fapar) %>% 
    mutate(doy = yday(date)) %>% 
    group_by(sitename, doy) %>% 
    summarise( evi_mean = mean( evi, na.rm=TRUE ), 
               evi_min  = min(  evi, na.rm=TRUE ), 
               evi_max  = max(  evi, na.rm=TRUE )
               ) %>%
    mutate( evi_min = ifelse( is.infinite(evi_min), NA, evi_min ), 
            evi_max = ifelse( is.infinite(evi_max), NA, evi_max )),
    by = c("sitename", "doy")
    )
  
# what's wrong with ...
# ddf_obs_dry %>% 
#   filter(sitename == "FR-LBr") %>% 
#   dplyr::select(date, fapar) %>% 
#   View()
#   ggplot(aes(x = date, y = fapar)) +
#   geom_line()

df_meandoy %>% 
  ggplot() +
  geom_line(aes(x = doy, y = fapar_mean), color = "black") +
  geom_ribbon(aes(x = doy, ymin = fapar_min, ymax = fapar_max), alpha=0.4) +
  geom_line(aes(x = doy, y = evi_mean), color = "green") +
  geom_ribbon(aes(x = doy, ymin = evi_min, ymax = evi_max), fill = "green", alpha=0.4) +
  geom_hline(yintercept = 0.15, color="red") +
  facet_wrap( ~sitename, ncol=4 ) +
  ylim(0,1)

ggsave("fig/fapar_seasonality_bysite.pdf", width=10, height = 18)
```


## AET

ET seasonality by site using AET (which included fAPAR in this simulation):

```{r  message=FALSE, warning=FALSE}
load("~/mct/data/out_eval_v2.RData")

ddf_modobs_dry <- out_eval$aet$fluxnet2015$data$meandoydf %>% 
  filter(sitename %in% df_flue_dry$site)

convert_le <- function(x){ x / (60*60*24)}

# ddf_modobs_dry %>%
#   mutate_at(vars(starts_with("mod_"), starts_with("obs_")), list(~convert_le)) %>%
#   ggplot(aes(x = doy)) +
#   geom_line(aes(y=obs_mean)) +
#   geom_ribbon(aes(ymin = obs_min, ymax = obs_max), alpha=0.4) +
#   geom_line(aes(y=mod_mean), color="red") +
#   geom_ribbon(aes(ymin = mod_min, ymax = mod_max), alpha=0.4, fill="tomato") +
#   facet_wrap( ~sitename, ncol=5 )
# 
# ggsave("fig/aet_seasonality_bysite.pdf", width=10, height = 10)
```

Modelled vs. observed ET using PET * fAPAR (red):
```{r  message=FALSE, warning=FALSE}
# load("~/mct/data/mod.RData")

## load v2 data - should be identical except for fapar (prescribed fapar not used for water balance in these simulations)
load("~/mct/data/mod_v2_EVI.RData") 
mod_evi <- mod
load("~/mct/data/mod_v2.RData")

ddf_modobs <- mod$daily %>% 
  bind_rows(.id="sitename") %>% 
  rename( fpar = fapar ) %>% 
  right_join(
    dplyr::select(ddf_obs_dry, sitename, date, latenth_obs = latenth, netrad_obs = netrad),
    by=c("sitename", "date")) %>%
  mutate(fpar_corr = (fpar - fpar_absmin)/(1.0-fpar_absmin)) %>% 
  mutate(fpar_corr = ifelse(fpar_corr<0, 0, fpar_corr)) %>% 
  mutate(pet = convert_le(pet), latenth_obs = convert_le(latenth_obs)) %>% 
  mutate(pet_fpar_corr = pet * fpar_corr) %>% 
  dplyr::select(sitename, date, fpar, fpar_corr, pet_fpar_corr, latenth_obs, netrad_obs)
  
ddf_modobs_evi <- mod_evi$daily %>% 
  bind_rows(.id="sitename") %>% 
  rename( evi = fapar ) %>% 
  right_join(
    dplyr::select(ddf_obs_dry, sitename, date, latenth_obs = latenth, netrad_obs = netrad),
    by=c("sitename", "date")) %>%
  mutate(evi_corr = (evi - evi_absmin)/(1.0-evi_absmin)) %>% 
  mutate(evi_corr = ifelse(evi_corr<0, 0, evi_corr)) %>% 
  mutate(pet = convert_le(pet)) %>% 
  mutate(pet_evi_corr = pet * evi_corr) %>% 
  dplyr::select(sitename, date, evi, evi_corr, pet_evi_corr)

ddf_modobs <- ddf_modobs %>% 
  left_join(ddf_modobs_evi, by = c("sitename", "date"))
    
out_modobs_fpar <- analyse_modobs2(ddf_modobs, "pet_fpar_corr", "latenth_obs", type = "heat")
out_modobs_fpar$gg +
  labs(x = "Modelled PET * FPAR (W m-2)", y = "Observed ET (W m-2)")
ggsave("fig/modobs_latenth_pet_fpar_corr.pdf", width=6, height = 6)

out_modobs_evi <- analyse_modobs2(ddf_modobs, "pet_evi_corr", "latenth_obs", type = "heat")
out_modobs_evi$gg +
  labs(x = "Modelled PET * EVI (W m-2)", y = "Observed ET (W m-2)")
ggsave("fig/modobs_latenth_pet_evi_corr.pdf", width=6, height = 6)
```

## PET * fAPAR

Modelled vs. observed ET seasonality by site using PET (dashed red, not shown), AET (blue), and PET * fAPAR (red):
```{r}
ddf_modobs_meandoy <- ddf_modobs %>% 
  dplyr::mutate(doy = yday(date)) %>% 
  # dplyr::mutate(latenth_obs = convert_le(latenth_obs)) %>% 
  dplyr::group_by(sitename, doy) %>% 
  dplyr::summarise_at( 
    vars(starts_with("fapar_"), starts_with("pet"), starts_with("latenth")), 
    list(~mean, ~min, ~max), na.rm=TRUE ) %>% 
  dplyr::mutate_at( vars(-group_cols()), ~ifelse(is.infinite(.), NA, .))
  # dplyr::left_join(
  #     mutate_at( 
  #       dplyr::select(ddf_modobs_dry, sitename, doy, starts_with("mod_")), 
  #       vars(starts_with("mod_"), starts_with("obs_")), 
  #       list(~convert_le)),
  #     by=c("sitename", "doy")
  #     )

# # plot fapar "corrected"
# ddf_mod_meandoy %>% 
#   ggplot(aes(x = doy)) +
#   geom_line(aes(y=fapar_corr_mean)) +
#   geom_ribbon(aes(ymin = fapar_corr_min, ymax = fapar_corr_max), alpha=0.4) +
#   facet_wrap( ~sitename, ncol=5 )

#ggsave("fig/fapar_corr_seasonality_bysite.pdf", width=10, height = 10)

# plot latent heat
ddf_modobs_meandoy %>% 
  ggplot(aes(x = doy)) +
  geom_line(aes(y=latenth_obs_mean)) +
  geom_ribbon(aes(ymin = latenth_obs_min, ymax = latenth_obs_max), alpha=0.4) +
  geom_line(aes(y=pet_fpar_corr_mean), color="red") +
  geom_line(aes(y=pet_evi_corr_mean), color="green") +
  geom_ribbon(aes(ymin = pet_fpar_corr_min, ymax = pet_fpar_corr_max), alpha=0.4, fill="red") +
  geom_ribbon(aes(ymin = pet_evi_corr_min, ymax = pet_evi_corr_max), alpha=0.4, fill="green") +
  # geom_line(aes(y=mod_mean), color="dodgerblue") +
  # geom_ribbon(aes(ymin = mod_min, ymax = mod_max), alpha=0.4, fill="dodgerblue") +
  # geom_line(aes(y=pet_mean), color="red", linetype="dashed") +
  facet_wrap( ~sitename, ncol=4 )

ggsave("fig/et_pet_fapar_corr_seasonality_bysite.pdf", width=10, height = 10)
```


