---
title: "Rooting depth by site (Schenk & Jackson, 2002)"
author: "Beni Stocker"
date: "4/30/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---
  
```{r setup, include=FALSE}
library(dplyr)
library(rbeni)
library(tidyr)
library(purrr)
library(ncdf4)
library(lubridate)
library(extRemes)
library(R.utils)
library(readr)
library(ggplot2)

source("R/mct2.R")
source("R/get_plantwhc_mct_bysite.R")
source("R/get_plantwhc_mct_simsuite.R")
source("R/calc_soilparams.R")
source("R/calc_zroot.R")
source("R/extract_return_level.R")
dir_climate <- "/Users/benjaminstocker/sofun/output_nc_global_sofun"
dir_fapar <- "~/sofun/input_sj02zroot_sofun/sitedata/fapar/"
```


## Prepare data
Get meta info of sites (lon, lat)
```{r}
df_sites <- read_csv("~/data/rootingdepth/root_profiles_schenkjackson02/data/root_profiles_D50D95.csv") %>%
  dplyr::filter(Wetland == "N" & Anthropogenic == "N" & Schenk_Jackson_2002 == "YES") %>% 
  dplyr::rename(sitename = ID, lat = Latitude, lon = Longitude) %>% 
  dplyr::mutate(elv = ifelse(elv==-999, NA, elv)) %>% 
  dplyr::filter(lon!=-999 & lat!=-999) %>% 
  dplyr::mutate(year_start = 1982, year_end = 2011) %>% 
  dplyr::select(sitename, lon, lat, elv, year_start, year_end)
```

Extract all the climate data (sofun output)
```{r eval=FALSE}
df <- tibble( year = 1982:2011 ) %>% 
  rowwise() %>%
  dplyr::mutate(filn_pet  = paste0( dir_climate, "/s1_fapar3g_v4_global.d.pet",  year, ".nc"),
                filn_wbal = paste0( dir_climate, "/s1_fapar3g_v4_global.d.wbal", year, ".nc")) %>%
  dplyr::mutate(data_pet  = purrr::map(filn_pet,  ~extract_pointdata_allsites(., df_sites, get_time = TRUE)),
                data_wbal = purrr::map(filn_wbal, ~extract_pointdata_allsites(., df_sites, get_time = TRUE)))
```

Re-arrange data to a flat table
```{r eval=FALSE}
ddf <- df %>% 
  dplyr::select(-filn_pet, -filn_wbal) %>% 
  tidyr::unnest(data_pet) %>% 
  dplyr::select(-year_start, -year_end, -year) %>% 
  tidyr::unnest(data) %>% 
  dplyr::rename(pet = V1) %>% 
  left_join(
    df %>% 
      dplyr::select(-filn_pet, -filn_wbal) %>% 
      tidyr::unnest(data_wbal) %>% 
      dplyr::select(-year_start, -year_end, -year) %>% 
      tidyr::unnest(data) %>% 
      dplyr::rename(water_to_soil = V1),
    by = c("lon", "lat", "sitename", "elv", "date")
  ) %>% 
  dplyr::select(-lon, -lat, -elv)

```

Read all the fapar data (interpolated to daily)
```{r eval=FALSE}
myread_csv <- function(filn){
  if (file.exists(filn)){
    df <- read_csv(filn) %>% 
      dplyr::select(date, fapar = modisvar_interpol )
  } else {
    df <- tibble(date = NA, fapar = NA)
  }
  return(df)
}
ddf <- df_sites %>% 
  # slice(1:10) %>%
  dplyr::mutate( filn = paste0( dir_fapar, "/", sitename, "dfapar_MODIS_FPAR_MCD15A3H_gee_MCD15A3H_", sitename, "_gee_subset.csv" ) ) %>% 
  dplyr::mutate( data_fapar = purrr::map(filn, ~myread_csv(.)) ) %>% 
  tidyr::unnest( data_fapar ) %>% 
  tidyr::drop_na(fapar) %>% 
  dplyr::select(sitename, date, fapar) %>% 
  dplyr::right_join(ddf, by = c("sitename", "date"))
```


Set fapar to zero where pet is zero (arctic night?)
```{r eval=FALSE}
ddf <- ddf %>%
  rowwise() %>% 
  dplyr::mutate(fapar = ifelse(pet==0 & is.na(fapar), 0, fapar))

# library(ggplot2)
# ddf %>% 
#   dplyr::filter(lubridate::year(date) %in% 2000:2005) %>% 
#   ggplot(aes(x=date, y=wbal - fapar * pet)) +
#   geom_line()

```

Add daytime VPD and calculate VPD-stress functions.
```{r}
f_vpd_oren01 <- function(q, D){
  q/(q+a) * (b + m *ln(D))
}
f_vpd_bigleaf <- function(D){
  g0 + 1.6 * (1 + g1/sqrt(D))
}
bigleaf::stomatal.sensitivity
```


Calculate wbal and nest data frames per site
```{r eval=FALSE}
ddf <- ddf %>% 
  dplyr::mutate(wbal = water_to_soil - fapar * pet) %>% 
  tidyr::drop_na(wbal) %>% 
  group_by(sitename) %>% 
  tidyr::nest() 
```

Remove sites where not the full time series is available
```{r eval=FALSE}
ddf <- ddf %>%
  dplyr::mutate(len = purrr::map_int(data, ~nrow(.))) %>% 
  dplyr::filter(len == 10948)
```


## Apply the MCT function

Apply the MCT function in different configurations to get plant-adjusted WHC across the rooting zone WCH$^\ast$.
```{r eval=FALSE}
ddf <- ddf %>% 
  dplyr::mutate( out_mct_50 = purrr::map(data, ~get_plantwhc_mct_bysite(., varname_wbal = "wbal", thresh_deficit = 0.5)) ) %>%
  dplyr::mutate( out_mct_75 = purrr::map(data, ~get_plantwhc_mct_bysite(., varname_wbal = "wbal", thresh_deficit = 0.25)) ) %>% 
  dplyr::mutate( out_mct_90 = purrr::map(data, ~get_plantwhc_mct_bysite(., varname_wbal = "wbal", thresh_deficit = 0.1)) ) %>% 
  dplyr::mutate( out_mct_95 = purrr::map(data, ~get_plantwhc_mct_bysite(., varname_wbal = "wbal", thresh_deficit = 0.05)) )
```

Write to file
```{r eval=FALSE}
save(ddf, file = "data/ddf_mct_simsuite.Rdata")
```

The following figure shows the distribution of WHC* values:
```{r}
load("data/ddf_mct_simsuite.Rdata")
ddf %>% 
  dplyr::select(sitename, out_mct_95) %>% 
  dplyr::mutate(whc_mct = purrr::map_dbl(out_mct_95, ~extract_return_level(., 10))) %>% 
  ggplot(aes(x = whc_mct, y = ..count..)) +
  geom_histogram(color = "black", alpha = 0.3, position="identity") +
  labs(title = "Plant rooting zone WHC*", subtitle = "10 y return period, 95% reduction of D", x = "WHC* (mm)")
```



## Extract soil texture information

Extract soil information from hi-res HWSD raster files (ideally should extract from shapefiles).
```{r}
source("R/extract_pointdata_allsites.R")
vars <- c("ROOTS", "T_SAND", "T_OC", "T_GRAVEL", "T_CLAY", "S_SAND", "S_OC", "S_GRAVEL", "S_CLAY")
df_lonlat <- df_sites %>% 
  distinct(lon, lat, .keep_all = TRUE) %>% 
  dplyr::select(lon, lat) %>% 
  dplyr::mutate(id = 1:n())
  
# extract by file
df_hwsd <- tibble( vars = vars ) %>% 
  dplyr::mutate(filn = paste0("~/data/soil/hwsd/hwsd_wieder/data/", vars, ".nc4")) %>% 
  dplyr::mutate(data = purrr::map(filn, ~extract_pointdata_allsites(., df_lonlat, get_time = FALSE)))

# re-arrange data
df_sites <- df_hwsd %>% 
  tidyr::unnest(data) %>% 
  dplyr::mutate(data = purrr::map_dbl(data, ~unlist(.))) %>% 
  dplyr::select(-filn, -lon, -lat) %>% 
  tidyr::spread(vars, data) %>% 
  dplyr::right_join(df_lonlat, by = "id") %>% 
  dplyr::right_join(df_sites, by = c("lon", "lat"))
```

Calculate FC, PWP, and WHC from texture data.
```{r}
# top-soil
df_sites_topsoil <- df_sites %>%
  dplyr::select(sitename, roots = ROOTS, fclay = T_CLAY, fgravel = T_GRAVEL, forg = T_OC, fsand = T_SAND) %>% 
  dplyr::mutate(fclay   = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, fclay),
                fgravel = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, fgravel),
                forg    = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, forg),
                fsand   = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, fsand)) %>% 
  calc_soilparams(method = "balland")

df_sites_topsoil_mean <- df_sites_topsoil %>% 
  dplyr::select(roots, whc) %>% 
  dplyr::group_by() %>% 
  dplyr::summarise_all(list( ~mean(., na.rm=TRUE)))

df_sites_topsoil <- df_sites_topsoil %>% 
  dplyr::mutate( whc = ifelse(is.na(whc), df_sites_topsoil_mean$whc, whc))

# sub-soil
df_sites_subsoil <- df_sites %>%
  dplyr::select(sitename, roots = ROOTS, fclay = S_CLAY, fgravel = S_GRAVEL, forg = S_OC, fsand = S_SAND) %>% 
  dplyr::mutate(fclay   = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, fclay),
                fgravel = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, fgravel),
                forg    = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, forg),
                fsand   = ifelse(fclay==0 & fgravel ==0 & fsand ==0, NA, fsand)) %>% 
  calc_soilparams(method = "balland")

df_sites_subsoil_mean <- df_sites_subsoil %>% 
  dplyr::select(roots, whc) %>% 
  dplyr::group_by() %>% 
  dplyr::summarise_all(list( ~mean(., na.rm=TRUE)))

df_sites_subsoil <- df_sites_subsoil %>% 
  dplyr::mutate( whc = ifelse(is.na(whc), df_sites_subsoil_mean$whc, whc))
```

Plot the distribution of values.
```{r}
df_sites_topsoil %>% 
  ggplot(aes(x = whc, y = ..count..)) +
  geom_histogram(color = "black", alpha = 0.3, position="identity") +
  labs(title = "Top soil WHC", subtitle = " based on HWSD soil texture data", x = "WHC (m3/m3)")

df_sites_subsoil %>% 
  ggplot(aes(x = whc, y = ..count..)) +
  geom_histogram(color = "black", alpha = 0.3, position="identity") +
  labs(title = "Sub soil WHC", subtitle = " based on HWSD soil texture data", x = "WHC (m3/m3)")
```

## Calculate rooting depth.

```{r}
ddf_20y_90t <- ddf %>%
  dplyr::left_join(dplyr::select(df_sites_subsoil, sitename, whc_s = whc, roots), by = "sitename") %>% 
  dplyr::left_join(dplyr::select(df_sites_topsoil, sitename, whc_t = whc), by = "sitename") %>% 
  dplyr::mutate(whc_mct = purrr::map_dbl(out_mct_90, ~extract_return_level(., 20))) %>%
  dplyr::select(sitename, whc_mct, whc_t, whc_s, roots) %>% 
  rowwise() %>% 
  dplyr::mutate(zroot = calc_zroot(whc_mct, whc_t, whc_s, roots))

ddf_10y_95t <- ddf %>%
  dplyr::left_join(dplyr::select(df_sites_subsoil, sitename, whc_s = whc, roots), by = "sitename") %>% 
  dplyr::left_join(dplyr::select(df_sites_topsoil, sitename, whc_t = whc), by = "sitename") %>% 
  dplyr::mutate(whc_mct = purrr::map_dbl(out_mct_95, ~extract_return_level(., 10))) %>%
  dplyr::select(sitename, whc_mct, whc_t, whc_s, roots) %>% 
  rowwise() %>% 
  dplyr::mutate(zroot = calc_zroot(whc_mct, whc_t, whc_s, roots))

ddf_05y_95t <- ddf %>%
  dplyr::left_join(dplyr::select(df_sites_subsoil, sitename, whc_s = whc, roots), by = "sitename") %>% 
  dplyr::left_join(dplyr::select(df_sites_topsoil, sitename, whc_t = whc), by = "sitename") %>% 
  dplyr::mutate(whc_mct = purrr::map_dbl(out_mct_95, ~extract_return_level(., 5))) %>%
  dplyr::select(sitename, whc_mct, whc_t, whc_s, roots) %>% 
  rowwise() %>% 
  dplyr::mutate(zroot = calc_zroot(whc_mct, whc_t, whc_s, roots))
```

The following figure shows the distribution of zroot* values:
```{r}
ddf_05y_95t %>% 
  ggplot(aes(x = zroot, y = ..count..)) +
  geom_histogram(color = "black", alpha = 0.3, position="identity") +
  labs(title = "Plant rooting depth", subtitle = "10 y return period, 95% reduction of D", x = "zroot* (mm)")
```

## Comparison to observations

Combine data frames.
```{r}
df_modobs <- read_csv("/alphadata01/bstocker/data/rootingdepth/root_profiles_schenkjackson02/data/root_profiles_D50D95.csv") %>% 
  dplyr::filter(Wetland == "N" & Anthropogenic == "N" & Schenk_Jackson_2002 == "YES") %>% 
  dplyr::rename(sitename = ID) %>% 
  dplyr::left_join(dplyr::select(ddf_05y_95t, zroot_05y_95t = zroot, sitename), by = "sitename") %>% 
  dplyr::left_join(dplyr::select(ddf_10y_95t, zroot_10y_95t = zroot, sitename), by = "sitename") %>% 
  dplyr::left_join(dplyr::select(ddf_20y_90t, zroot_20y_90t = zroot, sitename), by = "sitename") %>% 
  dplyr::mutate(D50 = 1000 * D50, D95 = 1000 * D95, D50_extrapolated = 1000 * D50_extrapolated, D95_extrapolated = 1000 * D95_extrapolated)
```

### Distribution of values
```{r}
df_modobs %>% 
  dplyr::select(sitename, Vegetation, obs = D95_extrapolated, mod = zroot_05y_95t) %>% 
  tidyr::gather(key = "source", value = "zroot", c(mod, obs)) %>% 
  ggplot() +
  geom_histogram(
    aes(x = zroot, y = ..count.., fill = source), 
    color = "black", alpha = 0.3, position="identity") +
  scale_fill_manual(name = "", values = c("black", "red")) +
  labs(title = "Distribution of rooting depth (mm), v3", x = "Rooting depth (mm)")
ggsave("fig/hist_zroot_modobs.pdf", height = 4, width = 6)
```

### Comparison by vegetation type.
```{r}
df_modobs %>% 
  dplyr::select(sitename, Vegetation, obs = D95_extrapolated, mod = zroot_05y_95t) %>% 
  tidyr::gather(key = "source", value = "zroot", c(mod, obs)) %>% 
  # dplyr::filter(source == "obs") %>% 
  ggplot() +
  geom_boxplot(aes(x = Vegetation, y = -zroot, fill = source)) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Observed and modelled by vegetation type", subtitle = "Obs.: 95% quantile  Mod.: 5-yr return period, 5% deficit reduction threshold", y = "Rooting depth (mm)")
ggsave("fig/modobs_boxplot_zroot_95_5y.pdf", width = 9, height = 6)

df_modobs %>% 
  dplyr::select(sitename, Vegetation, obs = D95_extrapolated, mod = zroot_20y_90t) %>% 
  tidyr::gather(key = "source", value = "zroot", c(mod, obs)) %>% 
  # dplyr::filter(source == "obs") %>% 
  ggplot() +
  geom_boxplot(aes(x = Vegetation, y = -zroot, fill = source)) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Observed and modelled by vegetation type", subtitle = "Obs.: 95% quantile  Mod.: 20-yr return period, 10% deficit reduction threshold", y = "Rooting depth (mm)")
```

### Comparison site by site.
```{r}
out <- df_modobs %>% 
  analyse_modobs2(
    mod = "zroot_05y_95t", 
    obs = "D95_extrapolated"
    )
out$gg +  
  labs(
    title = "Obs.: 95% quantile  Mod.: 5-yr return period, 5% deficit reduction threshold", 
    x = "Modelled rooting depth (mm)", 
    y = "Observed rooting depth (mm)"
    )
```
