---
title: "Trait gradient analysis of rooting depth"
author: "Beni Stocker"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(rbeni)
library(raster)
library(tibble)
library(ingestr)
library(ggplot2)
library(ggridges)
library(ggrepel)
library(cowplot)
```

The data is downloaded from the google spreadsheet [RSIP Working Copy], tab 'Analysis sheet', into a CSV (19.9.2019) and saved as `data/RSIP_Analysis_sheet.csv`.

Use `gsheet::gsheet2tbl`

Read the data.
```{r}
# df <- read_csv("~/data/rootingdepth/rsip/RSIP_Analysis_sheet_210409.csv") %>%      # previously done with "data/RSIP_Analysis_sheet.csv"
df <- read_csv("~/data/rootingdepth/rsip/RSIP_Analysis_sheet_210721.csv") %>%
  rename(lon = Long, lat = Lat) %>% 
  rowid_to_column(var = "id") %>% 
  
  ## problem: some have a reference error
  dplyr::filter(lon != "#REF!") %>% 
  mutate(lon = as.numeric(lon), lat = as.numeric(lat), 
         Dr = as.numeric(Dr),
         wtd = as.numeric(Water_Table_Depth_Fan))
  
print(paste("Total number of entries:", nrow(df)))
```


## Trait gradient analysis

First, remove data from herbaceous plants that occurr at sites together with woody plants. Group grasses and forbs into "herbaceous" and trees, shrubs, and semi-shrubs into "woody".
```{r}
df2 <- df %>% 
  mutate(sitename = paste0("i_", as.character(lon), "_", as.character(lat))) %>% 
  mutate(type = ifelse(Growth_form %in% c("Forb", "Grass"), "herb",
                       ifelse(Growth_form %in% c("Tree", "Shrub", "Semi-shrub"), "woody", NA))) %>% 
  filter(!is.na(type))

find_coexisting <- function(vec){
  vec <- unique(vec)
  if (length(vec) > 1){
    return(TRUE)
  } else {
    return(FALSE)
  }
}

tmp <- df2 %>% 
  group_by(sitename) %>% 
  summarise(coexisting = find_coexisting(type))

df3 <- df2 %>% 
  left_join(tmp, by = "sitename") %>% 
  filter(!(type == "herb" & coexisting))
```

## Plot depth heigh relationship

Remove the significant effect of shoot height. Note that 3230 data points have missing Hs. The filtering here reduces the dataset to 2287 entries. 
```{r}
df4 <- df3 %>% 
  mutate(Hs = as.numeric(Hs)) %>% 
  filter(Hs > 0 & Dr > 0)

df4 %>% 
  ggplot(aes(Hs, Dr)) +
  geom_point(alpha = 0.5) +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x = expression(italic("H")[s] ~ "(m)"),
       y = expression(italic("z")[r] ~ "(m)"))

ggsave("fig/height_depth.pdf", width = 6, height = 5)
ggsave("fig/height_depth.png", width = 6, height = 5)
```

```{r}
linmod <- lm(log(Dr) ~ log(Hs), 
             data = df4, 
             na.action = "na.omit")
```

```{r}
plot(linmod)
hist(linmod$residuals)
```
Consider residuals for the TGA.
```{r}
df4 <- df4 %>% 
  mutate(Dr_res = linmod$residuals)
```


<!-- Consider Hs-normalised values -->
<!-- ```{r} -->
<!-- df %>%  -->
<!--   mutate(Dr_norm = Dr / Hs^linmod$coefficients["log(Hs)"]) %>%  -->
<!--   ggplot(aes(Hs, Dr_norm)) + -->
<!--   geom_point(alpha = 0.5) + -->
<!--   geom_smooth(method = "lm") -->

<!-- df %>%  -->
<!--   ggplot(aes(Hs, Dr)) + -->
<!--   geom_point(alpha = 0.5) + -->
<!--   geom_smooth(method = "lm") -->
<!-- ``` -->

## Aggregate to sites

```{r}
df_sites2 <- df4 %>% 
  mutate(sitename = paste0("i_", as.character(lon), "_", as.character(lat))) %>% 
  group_by(sitename) %>% 
  summarise(Dr = mean(Dr, na.rm = TRUE),
            Dr_res = mean(Dr_res, na.rm = TRUE))

## add WWF biome info
df_wwf_sites2 <- ingest(
  df4 %>% 
    distinct(lon, lat) %>% 
    mutate(sitename = paste0("i_", as.character(lon), "_", as.character(lat))),
  source = "wwf",
  dir = "~/data/biomes/wwf_ecoregions/official/",
  settings = list(layer = "wwf_terr_ecos")
  )%>% 
  mutate(data = purrr::map(data, ~slice(., 1))) %>% 
  unnest(data)

df_sites2 <- df_sites2 %>% 
  left_join(
    df_wwf_sites2 %>% 
      dplyr::select(sitename, biome = BIOME, biome_name = BIOME_NAME),
    by = "sitename"
  )

df_sites2 %>%
  separate(sitename, into = c(NA, "lon", "lat"), sep = "_") %>% 
  mutate(lon = as.numeric(lon), lat = as.numeric(lat)) %>% 
  write_csv(file = "data/df_sites_rsip_tga.csv")
```

Add site mean to full data
```{r}
df4 <- df_sites2 %>% 
  dplyr::select(sitename, Dr_sitemean = Dr, Dr_res_sitemean = Dr_res) %>% 
  right_join(
    df4 %>% 
      mutate(sitename = paste0("i_", as.character(lon), "_", as.character(lat))), 
    by = "sitename"
    )
```


## Filter data

Use data only for sites where at least 3 data points are available. Reduces data from 1,497 to 1,197 points.
```{r}
use_sites <- df4 %>% 
  dplyr::select(sitename, Species) %>% 
  group_by(sitename) %>% 
  summarise(n = n()) %>% 
  dplyr::filter(n >= 3) %>% 
  pull(sitename)

df5 <- df4 %>% 
  dplyr::filter(sitename %in% use_sites)
```

Use data only for species that appear in at least 3 sites => 35 species and 166 data points.
```{r}
use_species <- df5 %>%
  filter(Species != "NA") %>% 
  dplyr::select(sitename, Species) %>% 
  distinct() %>% 
  group_by(Species) %>% 
  summarise(n = n()) %>% 
  dplyr::filter(n >= 3) %>% 
  pull(Species)

df6 <- df5 %>% 
  dplyr::filter(Species %in% use_species)
```

```{r}
# test : number of species per site
df6 %>% 
  dplyr::select(sitename, Species) %>% 
  distinct() %>% 
  group_by(sitename) %>% 
  summarise(n = n())

# test : number of sites per species
df6 %>% 
  dplyr::select(sitename, Species) %>% 
  distinct() %>% 
  group_by(Species) %>% 
  summarise(n = n()) %>% 
  arrange(n)

saveRDS(df6, file = "data/df_tga.rds")
```

## TGA

Plot.
```{r eval=FALSE}
df6 <- readRDS("data/df_tga.rds")

df6 %>% 
  ggplot(aes(x = Dr_res_sitemean, y = Dr_res)) +  # , color = Species
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")
```

Plot just the lines
```{r}
gg1 <- df6 %>% 
  group_by(Family) %>% 
  ggplot(aes(x = Dr_res_sitemean, y = Dr_res, group = Species, color = Family)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.5, alpha = 0.2) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  theme_classic() +
  geom_point(alpha = 0.3) +
  labs(x = expression("Site mean ln" ~ italic("z")[r] ~ "(unitless)"),
       y = expression("ln" ~ italic("z")[r] ~ "(unitless)")) +
  theme(legend.title = element_blank())

gg1
ggsave("fig/tga_rsip.pdf", width = 6, height = 4)
ggsave("fig/tga_rsip.png", width = 6, height = 4)
```


<!-- Fit linear regressions by species for Dr (not residual). -->
<!-- ```{r} -->
<!-- get_width <- function(df){ -->
<!--   df %>% pull(Dr) %>% range() %>% diff() -->
<!-- } -->

<!-- df_tga <- df5 %>%  -->
<!--   dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>%  -->
<!--   group_by(Species) %>%  -->
<!--   nest() %>%  -->
<!--   mutate(linmod = purrr::map(data, ~lm(Dr ~ Dr_sitemean, data = .)), -->
<!--          width = purrr::map_dbl(data, ~get_width(.))) %>%  -->
<!--   mutate(slope = purrr::map_dbl(linmod, ~coef(.)[2])) %>%  -->
<!--   left_join(df5 %>%  -->
<!--               dplyr::select(Species, Family) %>%  -->
<!--               distinct(),  -->
<!--             by = "Species") -->

<!-- df_tga %>%  -->
<!--   ggplot() + -->
<!--   geom_histogram(aes(slope, ..density..), fill = "royalblue", binwidth = 0.4, alpha = 0.5) + -->
<!--   geom_density(aes(slope, ..density..), color = "royalblue") + -->
<!--   xlim(-2,3) -->

<!-- df_tga %>%  -->
<!--   ggplot(aes(width, slope)) + -->
<!--   geom_point() + -->
<!--   geom_smooth(method = "lm") + -->
<!--   geom_hline(yintercept = 1, linetype = "dotted") -->

<!-- linmod2 <- lm(slope ~ width, data = df_tga) -->
<!-- summary(linmod2) -->

<!-- df_tga %>%  -->
<!--   group_by(Family) %>%  -->
<!--   summarise(slope = mean(slope)) %>%  -->
<!--   mutate(Family = forcats::fct_reorder(Family, slope)) %>%  -->
<!--   drop_na() %>%  -->
<!--   ggplot(aes(Family, slope)) + -->
<!--   geom_bar(stat = "identity") + -->
<!--   coord_flip() -->
<!-- ``` -->

Fit linear regressions by species for `Dr_res` (residual of Dr from model log(Dr) ~ log(Hs)).
```{r}
get_width_res <- function(df){
  df %>% pull(Dr_res) %>% range() %>% diff()
}

df_tga_res <- df5 %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  nest() %>% 
  mutate(linmod = purrr::map(data, ~lm(Dr_res ~ Dr_res_sitemean, data = .)),
         width = purrr::map_dbl(data, ~get_width_res(.))) %>% 
  mutate(slope = purrr::map_dbl(linmod, ~coef(.)[2])) %>% 
  left_join(df5 %>% 
              dplyr::select(Species, Family) %>% 
              distinct(), 
            by = "Species") %>% 
  
  ## remove outlier slope
  filter(slope < 30) %>% 
  mutate(n = purrr::map_int(data, ~nrow(.)))

gg2 <- df_tga_res %>% 
  ggplot() +
  geom_histogram(aes(slope, ..count..), binwidth = 0.4, alpha = 0.7) +
  # geom_density(aes(slope, ..density..), color = "red") +
  geom_vline(xintercept = 1.0, linetype = "dotted") +
  theme_classic() +
  labs(x = "Slope", y = "Count")


df_tga_res %>% 
  ggplot(aes(width, slope)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 1, linetype = "dotted")

df_tga_res %>% 
  ggplot(aes(width, abs(slope))) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 1, linetype = "dotted")

linmod2 <- lm(slope ~ width, data = df_tga_res)
summary(linmod2)

## species-level
gg3 <- df_tga_res %>% 
  drop_na() %>% 
  ggplot(aes(forcats::fct_reorder(Species, slope), slope)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1.0, linetype = "dotted") +
  coord_flip() +
  labs(y = "Slope", x = "")

## family level
df_tga_res %>% 
  group_by(Family) %>% 
  summarise(slope = mean(slope), n = n()) %>% 
  mutate(family_n = paste0(Family, " (", as.character(n), ")")) %>% 
  mutate(family_n = forcats::fct_reorder(family_n, slope)) %>% 
  drop_na() %>% 
  ggplot(aes(family_n, slope)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1.0, linetype = "dotted") +
  coord_flip()
```

## Publication figures

```{r}
# bottom_row <- plot_grid(gg2, gg3, ncol = 4, labels = c('b', 'c'), rel_widths = c(0.3, 0.7))
# plot_grid(gg1, bottom_row, ncol = 1, labels = c('a', ''))

top_row <- plot_grid(gg1, gg2, ncol = 2, labels = c('a', 'b'), rel_widths = c(0.7, 0.3))
plot_grid(top_row, gg3, nrow = 2, labels = c('', 'c'), rel_heights = c(1, 1.5) )

ggsave("fig/tga.pdf", width = 12, height = 9)
ggsave("fig/tga.png", width = 12, height = 9)
```
