---
title: "RSIP data analysis"
author: "Beni Stocker"
date: "9/19/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(rbeni)
library(raster)
library(tibble)
```

The data is downloaded from the google spreadsheet [RSIP Working Copy](https://docs.google.com/spreadsheets/d/1IW7CiLaw2xEr3MYuLwZ5mJWEbhAEU_hP9J6od5vb-jk/edit#gid=1200594845), tab 'Analysis sheet', into a CSV (19.9.2019) and saved as `data/RSIP_Analysis_sheet.csv`.

Read the data.
```{r}
df <- read_csv("data/RSIP_Analysis_sheet.csv") %>% 
  rename(lon = Long, lat = Lat) %>% 
  rowid_to_column(var = "id")
print(paste("Total number of entries:", nrow(df)))
```

Bin the data into grid points, summarise, and expand to grid
```{r}
## bin
dlon <- 5.0
dlat <- 5.0
lon_breaks <- seq(from = -180, to = 180, by = dlon)
lat_breaks <- seq(from = -90, to = 90, by = dlat)

df <- df %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    # labels = as.character(seq(length(lon_breaks)-1))
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    # labels = as.character(seq(length(lat_breaks)-1))
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2)

#View(select(df, lon, lat, ilon, ilat, lon_lower, lon_upper, lat_lower, lat_upper, lon_mid, lat_mid))

df <- df %>% 
  group_by(lon_mid, lat_mid) %>% 
  drop_na(lon_mid, lat_mid)
```

Summarise
```{r}
df_agg <- df %>% 
  summarise(mean_Dr = mean(Dr, na.rm = TRUE)) %>% 
  left_join(
    summarise(df, median_Dr = median(Dr, na.rm = TRUE)),
    by = c("lon_mid", "lat_mid")
  ) %>% 
  left_join(
    summarise(df, max_Dr = max(Dr, na.rm = TRUE)),
    by = c("lon_mid", "lat_mid")
  ) %>% 
  left_join(
    summarise(df, q75 = quantile(Dr, probs = 0.75, na.rm = TRUE) ),
    by = c("lon_mid", "lat_mid")
  ) %>% 
  left_join(
    summarise(df, q90 = quantile(Dr, probs = 0.90, na.rm = TRUE) ),
    by = c("lon_mid", "lat_mid")
  ) %>% 
  left_join(
    summarise(df, q95 = quantile(Dr, probs = 0.95, na.rm = TRUE) ),
    by = c("lon_mid", "lat_mid")
  ) %>% 
  left_join(
    summarise(df, n_Dr = n()),
    by = c("lon_mid", "lat_mid")
  )
df <- df %>% ungroup()
```


## Plots

### Distribution of values by individual data points.
```{r}
df %>% 
  ggplot(aes(x = Dr)) +
  geom_histogram(binwidth = 1, fill = "grey70", colour = "black") +
  lims(x = c(0,125))
```


### Distribution of values by bins.
```{r}
df_agg %>% 
  ggplot(aes(x = max_Dr)) +
  geom_histogram(binwidth = 2, fill = "grey70", colour = "black")
```

### Global map
```{r}
gg <- plot_map2(
  df_agg %>% 
    ungroup %>% 
    dplyr::filter(n_Dr > 2) %>% 
    dplyr::rename(lon = lon_mid, lat = lat_mid) %>%
    dplyr::select(lon, lat, median_Dr),
  breaks = seq(from = 0, to = 15, length.out = 11),
  maxval = 125
  )
gg
#ggsave("fig/test.pdf", width = 10, height = 8)
```

## Analyse by biome

This doesn't seem to work. For almost all the points given in `df` with their `lat` and `lon`, there is no biome information in the raster file. 

First, extract biome information from NCRS biome map.
```{r}
add_biome_ncrs <- function(df, path){

 rasta_biome <- raster::raster(path) %>% 
  subset(1, drop = TRUE)

  ## Extract data at points given by global model output (gridcells)
  df_biome_ncrs <- extract(rasta_biome, SpatialPoints(dplyr::select(df, lon, lat)), sp = TRUE) %>% 
    as_tibble()

  df <- df_biome_ncrs %>% 
    dplyr::select(-lat, -lon) %>% 
    bind_cols(df) %>% 
    rename(biome_ncrs = biomes) %>% 
    mutate(biome_ncrs = as.factor(biome_ncrs))

  biome_key <- tibble(
    biome_ncrs = as.factor(0:14),
    biome_ncrs_chr = c(
      "Ocean",
      "TUNDRA Permafrost",
      "TUNDRA Interfrost",
      "BOREAL Semi-arid",
      "BOREAL Humid",
      "TEMPERATE Semi-arid",
      "TEMPERATE Humid",
      "MEDITERRANEAN Warm",
      "MEDITERRANEAN Cold",
      "DESERT Tropical",
      "DESERT Temperate",
      "DESERT Cold",
      "TROPICAL Semi-arid",
      "TROPICAL Humid",
      "Ice"
      )
    )
  df <- df %>%
    left_join(biome_key, by = "biome_ncrs") 
  return(df)
}

df <- add_biome_ncrs(
  df, 
  path = "/alphadata01/bstocker/data/biomes/NCRS/global_biomes_geotiff/biomes.tif"
  )

df_agg <- add_biome_ncrs(
  df_agg %>% rename(lon = lon_mid, lat = lat_mid), 
  path = "/alphadata01/bstocker/data/biomes/NCRS/global_biomes_geotiff/biomes.tif"
  )
```

Plot data by biomes: Individual-level data
```{r}
df %>% 
  ggplot(aes(x = biome_ncrs_chr, y = -Dr)) +
  # geom_boxplot() +
  geom_violin() +
  theme(axis.text.x = element_text(angle=90, hjust=1))
```

## Add model results

```{r}
load("./data/df_20y_bl_v3.Rdata") # loading df_20y_bl_v3
df_20y_bl_v3 <- df_20y_bl_v3 %>% 
  add_biome_ncrs(path = "/alphadata01/bstocker/data/biomes/NCRS/global_biomes_geotiff/biomes.tif")

# df_20y_bl_v3 %>% 
#   ggplot(aes(x = biome_ncrs_chr, y = -zroot/10000)) +
#   # geom_boxplot() +
#   geom_violin() +
#   theme(axis.text.x = element_text(angle=90, hjust=1))

df %>% 
  dplyr::select(biome_ncrs_chr, zroot = Dr) %>% 
  mutate(source = "obs") %>% 
  bind_rows(
    df_20y_bl_v3 %>% 
      dplyr::select(biome_ncrs_chr, zroot) %>% 
      mutate(zroot = zroot/10000, source = "mod")
    ) %>% 
  ggplot(aes(x = biome_ncrs_chr, y = -zroot, fill = source, color = source)) +
  geom_boxplot() +
  # geom_violin() +
  # geom_bar(stat = "summary", fun.ymax = "max", position=position_dodge()) +
  theme(axis.text.x = element_text(angle=90, hjust=1))

df %>% 
  dplyr::select(biome_ncrs_chr, zroot = Dr) %>% 
  mutate(source = "obs") %>% 
  bind_rows(
    df_20y_bl_v3 %>% 
      dplyr::select(biome_ncrs_chr, zroot) %>% 
      mutate(zroot = zroot/10000, source = "mod")
    ) %>% 
  ggplot(aes(x = biome_ncrs_chr, fill = source, color = source)) +
  stat_summary_bin(aes(y = -zroot), fun.y = "min", geom = "bar", position=position_dodge()) +
  theme(axis.text.x = element_text(angle=90, hjust=1))
```