---
title: "rsofun at RSIP sites"
author: "Beni Stocker"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rsofun)
library(ingestr)
library(rbeni)
```

## Get sites

```{r cars}
df <- read_csv("~/data/rootingdepth/rsip/RSIP_Analysis_sheet_210409.csv") %>%      # previously done with "data/RSIP_Analysis_sheet.csv"
# df <- read_csv("~/data/rootingdepth/rsip/RSIP_Analysis_sheet_210721.csv") %>% 
  rename(lon = Long, lat = Lat) %>% 
  rowid_to_column(var = "id") %>% 
  
  ## problem: some have a reference error
  dplyr::filter(lon != "#REF!") %>% 
  mutate(lon = as.numeric(lon), lat = as.numeric(lat), 
         Dr = as.numeric(Dr),
         wtd = as.numeric(Water_Table_Depth_Fan))
  
print(paste("Total number of entries:", nrow(df)))
```

Add model results.
```{r}
df <- extract_nc(df %>% dplyr::select(id, lon, lat), "data/zroot_cwd80.nc") %>% 
  unnest(data) %>% 
  rename(zroot_cwdx80 = V1) %>% 
  mutate(zroot_cwdx80 = zroot_cwdx80 / 1000) %>% 
  right_join(df, by = c("lon", "lat", "id"))
```

1705 distinct sites based on lon and lat info.
```{r}
df_sites <- df %>% 
  mutate(sitename = paste0("i_", as.character(lon), "_", as.character(lat))) %>% 
  group_by(sitename) %>% 
  summarise(Dr = mean(Dr, na.rm = TRUE), zroot_cwdx80 = mean(zroot_cwdx80, na.rm = TRUE), wtd = mean(wtd, na.rm = TRUE))

## add WWF biome info
df_wwf_sites <- ingest(
  df %>% 
    distinct(lon, lat) %>% 
    mutate(sitename = paste0("i_", as.character(lon), "_", as.character(lat))),
  source = "wwf",
  dir = "~/data/biomes/wwf_ecoregions/official/",
  settings = list(layer = "wwf_terr_ecos")
  )%>% 
  mutate(data = purrr::map(data, ~slice(., 1))) %>% 
  unnest(data)

df_sites <- df_sites %>% 
  left_join(
    df_wwf_sites %>% 
      dplyr::select(sitename, biome = BIOME, biome_name = BIOME_NAME),
    by = "sitename"
  )

df_sites <- df_sites %>% 
  separate(sitename, c("prefix", "lon", "lat"), sep = "_") %>% 
  dplyr::select(-prefix) %>% 
  mutate(sitename = paste0("i_", 1:nrow(.))) %>% 
  mutate(lon = as.numeric(lon), lat = as.numeric(lat)) %>%
  mutate(
    year_start = 1989,
    year_end = 2018
  )

# add elevation data
if(grepl('eu-', Sys.info()['nodename'])){
  r <- raster::raster("~/data/etopo/ETOPO1_Bed_g_geotiff.tif")
}

# grab elev data
s <- st_as_sf(
  df,
  coords = c("lon","lat"),
  crs = 4326)
df$elv <- raster::extract(r, s)
df <- df %>%
  mutate(
    elv = ifelse(elv < 0, 0, elv)
  )

write_csv(df_sites, file = "data/df_sites_rsip.csv")  
```

## Get forcing data

Use `submit_forcing_rsip.sh` to run `rscript_forcing_rsip.R`. This saves files in subdirectory `"./data/forcing_rsip/"`.

## Test

Use `submit_rsofun_cwdx.sh` to run `rscript_rsofun_cwdx.R` for different S_0 (WHC) and diagnose it following the S_dEF method. This writes files into the directory `"./data/out_rsofun_cwdx/"`. 

```{r}
# ichunk <- 1
# use_whc <- 100

read_myfile <- function(ichunk, use_whc){
  path <- "./data/out_rsofun_cwdx/"
  filename <- file.path(path, paste0("out_rsofun_cwdx_whc_", as.character(use_whc), "_ichunk_", as.character(ichunk), ".rds"))
  if (file.exists(filename)){
    df <- readRDS(filename) %>% 
      dplyr::select(sitename, whc, s_def, s_cwdx80)
    return(df)
  } else {
    return(tibble())
  }
}

df <- purrr::map_dfr(
  as.list(seq(50)),
  ~read_myfile(., 100)
  ) %>% 
  mutate(setup = "whc_100") %>% 
  bind_rows(
    purrr::map_dfr(
      as.list(seq(50)),
      ~read_myfile(., 200)
      ) %>% 
      mutate(setup = "whc_200")
  ) %>% 
  bind_rows(
    purrr::map_dfr(
      as.list(seq(50)),
      ~read_myfile(., "NA")
      ) %>% 
      mutate(setup = "whc_NA")
  )

saveRDS(df, file = "data/df_test_cwdx_rsofun.rds")
df <- readRDS("data/df_test_cwdx_rsofun.rds")

df %>% 
  dplyr::filter(setup == "whc_NA") %>% 
  analyse_modobs2("whc", "s_def", type = "hex")

gg <- df %>% 
  dplyr::filter(setup == "whc_NA" & !is.nan(s_cwdx80) & !is.infinite(s_cwdx80) & s_cwdx80 < 1e4) %>%
  mutate(s_cwdx80 = remove_outliers(s_cwdx80)) %>% 
  analyse_modobs2("whc", "s_cwdx80", type = "hex")

gg$gg + xlim(0, 2000) + ylim(0, 2000)
```


