---
title: "Create rsofun-WHC forcing map"
author: "Beni Stocker"
date: "2024-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(terra)
library(rbeni)
```


Get WHC data for topsoil and subsoil per gridcell as data frame.
```{r}
rasta_whc_1m <- rast(here::here("data/whc_1m.nc"))
rasta_whc_2m <- rast(here::here("data/whc_2m.nc"))
```

Get vegetation height data and regrid to same grid as WHC.
```{r}
rasta_height <- rast(here::here("~/data/archive/vegheight_simard_2011/data/vegheight_simard_2011.nc"))
rasta_height_regridded <- resample(rasta_height, rasta_whc_1m, method = "bilinear")
```

Combine values of all rasters as data frame.
```{r}
df_map <- as.data.frame(rasta_whc_1m, xy = TRUE) |>
  as_tibble() |>
  rename(lon = x,
         lat = y) |>
  left_join(
    as.data.frame(rasta_whc_2m, xy = TRUE) |>
      as_tibble() |>
      rename(lon = x,
             lat = y),
      by = join_by(lon, lat)
    ) |>
  left_join(
    as.data.frame(rasta_height_regridded, xy = TRUE) |>
      as_tibble() |>
      rename(lon = x,
             lat = y,
             vegheight = Band1),
      by = join_by(lon, lat)
    )
```

Fit regression of rooting depth with tree height.
```{r}
df <- read_csv("~/data/rootingdepth/rsip/RSIP_Analysis_sheet_210721.csv") %>%
  rename(lon = Long, lat = Lat) %>% 
  rowid_to_column(var = "id") %>% 
  
  ## problem: some have a reference error
  dplyr::filter(lon != "#REF!") %>% 
  mutate(lon = as.numeric(lon), lat = as.numeric(lat), 
         Dr = as.numeric(Dr),
         wtd = as.numeric(Water_Table_Depth_Fan))
```

Explorations with tree height (`Hs`).
```{r}
df2 <- df |> 
  mutate(Hs = as.numeric(Hs)) |> 
  mutate(log_dr = log10(Dr), 
         log_hs = log10(Hs))|> 
  dplyr::filter(!is.nan(log_dr) & 
                  !is.nan(log_hs) & 
                  !is.na(log_dr) & 
                  !is.na(log_hs) & 
                  !is.infinite(log_dr) & 
                  !is.infinite(log_hs))

linmod <- lm(log_dr ~ log_hs, data = df2)

plot(linmod)
```

```{r}
# vegetation height in RSIP data is in m
df2 |> 
  ggplot(aes(Hs, after_stat(density))) +
  geom_histogram()

# vegetation height in map is also in m
df_map |> 
  ggplot(aes(vegheight, after_stat(density))) +
  geom_histogram()
```

Visualise relationship and quantile regression fits.
```{r}
df2 |> 
  ggplot(aes(log_hs, log_dr)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm") + 
  geom_quantile(color = "royalblue", alpha = 0.5,
                quantiles = seq(.1, .9, by = 0.1)) +
  geom_quantile(color = "tomato", quantiles = 0.1) +
  theme_classic()
```

Get the quantile regression model for the 10% quantile.
```{r}
qrmod <- quantreg::rq(
  log_dr ~ log_hs, 
  tau = 0.1,
  data = df2
  )
```

Predict 10% quantile of rooting depth as a function of vegetation height.
```{r}
df_map <- df_map |> 
  mutate(log_hs = log(vegheight))

df_map <- df_map |> 
  mutate(log_dr = predict(qrmod, newdata = df_map)) |> 
  mutate(dr = exp(log_dr))
```

Distribution of predicted rooting depths minima (10% quantiles).
```{r}
df_map |> 
  ggplot(aes(dr, after_stat(count))) +
  geom_histogram()
```

Convert rooting depth into root zone water storage capacity, using WHC info of top and subsoil.
```{r}
get_rzwsc <- function(dr, whc_1, whc_2){
  ifelse(dr <= 1,
         dr * whc_1,
         whc_1 + (dr - 1) * (whc_2/2)
         )
}

df_map <- df_map |> 
  
  # where no rooting depth estimate is available, assume the 25% quantile rooting depth
  # preferred over mean because missing rooting depth is where vegetation height is missing
  # which is mostly in non-forested vegetation.
  mutate(dr = ifelse(is.na(dr), quantile(df_map$dr, probs = 0.25, na.rm = TRUE), dr)) |> 
  
  # convert to storage capacity
  mutate(rzwsc = get_rzwsc(dr, whc_1m, whc_2m))
```

Distribution of predicted root zone water storage capacities.
```{r}
df_map |> 
  ggplot(aes(rzwsc, after_stat(count))) +
  geom_histogram() + 
  xlim(0, 1500)
```

Take maximum of S_CWDX80 and the rzwsc calculated above. And the same for rooting depth.
```{r}
df_scwdx80 <- rast(here::here("data/cwdx80.nc")) |> 
  as.data.frame(xy = TRUE) |> 
  as_tibble() |> 
  rename(lon = x, lat = y) |> 
  mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)) |>
  left_join(
    rast(here::here("data/zroot_cwd80.nc")) |> 
      as.data.frame(xy = TRUE) |> 
      as_tibble() |> 
      rename(lon = x, lat = y) |> 
      mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)),
    by = join_by(lon, lat)
  )

df_map <- df_map |> 
  mutate(lon = round(lon, digits = 3), lat = round(lat, digits = 3)) |>
  left_join(
    df_scwdx80,
    by = join_by(lon, lat)
  ) |> 
  mutate(
    cwdx80_forcing = ifelse(cwdx80 > rzwsc, cwdx80, rzwsc),
    zroot_cwdx80_forcing = ifelse(zroot_cwd80 > dr, zroot_cwd80, dr)
  )
```

Write to file
```{r}
nc <- df_to_grid(df_map, 
                 varnam = "cwdx80_forcing", 
                 lonnam = "lon", 
                 latnam = "lat"
                 )
write_nc2(nc, 
          varnams = "cwdx80_forcing", 
          lon = df_map$lon |> unique() |> sort(), 
          lat = df_map$lat |> unique() |> sort(), 
          path = here::here("data/cwdx80_forcing.nc"), 
          make_zdim = FALSE
          )

nc <- df_to_grid(df_map, 
                 varnam = "zroot_cwdx80_forcing", 
                 lonnam = "lon", 
                 latnam = "lat"
                 )
write_nc2(nc, 
          varnams = "zroot_cwdx80_forcing", 
          lon = df_map$lon |> unique() |> sort(), 
          lat = df_map$lat |> unique() |> sort(), 
          path = here::here("data/zroot_cwdx80_forcing.nc"), 
          make_zdim = FALSE
          )
```

Regrid to 0.1 deg.
```{r}
lon_breaks <- seq(from = floor(min(df_map$lon)), to = ceiling(max(df_map$lon)), by = 0.1)
lat_breaks <- seq(from = floor(min(df_map$lat)), to = ceiling(max(df_map$lat)), by = 0.1)

df_map <- df_map |>
  ungroup() |>
  mutate(ilon = cut(lon,
                    breaks = lon_breaks),
         ilat = cut(lat,
                    breaks = lat_breaks)) |>
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )) |>
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2) |>
  
  ## create cell name to associate with climate input
  dplyr::select(-ilon, -ilat, -lon_lower, -lon_upper, -lat_lower, -lat_upper)

df_map_agg <- df_map |> 
  group_by(lon_mid, lat_mid) |> 
  summarise(rzwsc = mean(rzwsc, na.rm = TRUE),
            cwdx80_forcing = mean(cwdx80_forcing, na.rm = TRUE)) |> 
  rename(lon = lon_mid, lat = lat_mid)
  # mutate(!!varnam := ifelse(is.nan(!!varnam), NA, !!varnam))
```

### Plot global WHC

```{r}
gg1 <- plot_map4(df_map_agg, 
                varnam = "rzwsc", 
                breaks = c(seq(0, 100, by = 20), 150, 200, 300, 500, 700, 900, 1200, Inf), 
                latmin = -60, latmax = 80,
                spacing = "constant", 
                maxval = 6000, 
                combine = TRUE, 
                colorscale = "batlowK", 
                legend_title = "(mm)",
                expand_size_y = 0.5,
                ocean = TRUE)
gg1

gg2 <- plot_map4(df_map_agg, 
                varnam = "cwdx80_forcing", 
                breaks = c(seq(0, 100, by = 20), 150, 200, 300, 500, 700, 900, 1200, Inf), 
                latmin = -60, latmax = 80,
                spacing = "constant", 
                maxval = 6000, 
                combine = TRUE, 
                colorscale = "batlowK", 
                legend_title = "(mm)",
                expand_size_y = 0.5,
                ocean = TRUE)
gg2
# ggsave("fig/map_whc_1m.pdf", width = 10, height = 5)
# ggsave("fig/map_whc_1m.png", width = 10, height = 5)
```