---
title: "Calculate rooting depth globally"
author: "Beni Stocker"
date: "4/30/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
library(rbeni)
library(dplyr)
library(ggplot2)
library(tidyr)
source("R/calc_zroot.R")
```

Get WHC based on soil texture
```{r echo=FALSE}
df_sr <- nc_to_df("/alphadata01/bstocker/data/soil/hwsd/hwsd_wieder/data_calculated_halfdeg/T_sr_WHC.nc", "whc") %>% 
  rename(whc_t = myvar) %>% 
  left_join(
    nc_to_df("/alphadata01/bstocker/data/soil/hwsd/hwsd_wieder/data_calculated_halfdeg/S_sr_WHC.nc", "whc") %>% 
      rename(whc_s = myvar),
    by = c("lon", "lat")
  ) %>% 
  drop_na()

df_bl <- nc_to_df("/alphadata01/bstocker/data/soil/hwsd/hwsd_wieder/data_calculated_halfdeg/T_bl_WHC.nc", "whc") %>% 
  rename(whc_t = myvar) %>% 
  left_join(
    nc_to_df("/alphadata01/bstocker/data/soil/hwsd/hwsd_wieder/data_calculated_halfdeg/S_bl_WHC.nc", "whc") %>% 
      rename(whc_s = myvar),
    by = c("lon", "lat")
  ) %>% 
  drop_na()
```

Add MCT-WHC with return period of 20 and 40 years.
```{r echo=FALSE}
# v1 Saxton & Rawls
df_sr_v1 <- df_sr %>% 
  left_join(
    nc_to_df("data/whc_mct_40y_v1.nc", "whc_mct_40y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_40y = myvar ),
    by = c("lon", "lat")
  )  %>% 
  left_join(
    nc_to_df("data/whc_mct_20y_v1.nc", "whc_mct_20y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_20y = myvar ),
    by = c("lon", "lat")
  )

# v3 Saxton & Rawls
df_sr_v3 <- df_sr %>% 
  left_join(
    nc_to_df("data/whc_mct_40y_v3.nc", "whc_mct_40y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_40y = myvar ),
    by = c("lon", "lat")
  )  %>% 
  left_join(
    nc_to_df("data/whc_mct_20y_v3.nc", "whc_mct_20y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_20y = myvar ),
    by = c("lon", "lat")
  )

# v1 Balland
df_bl_v1 <- df_bl %>% 
  left_join(
    nc_to_df("data/whc_mct_40y_v1.nc", "whc_mct_40y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_40y = myvar ),
    by = c("lon", "lat")
  )  %>% 
  left_join(
    nc_to_df("data/whc_mct_20y_v1.nc", "whc_mct_20y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_20y = myvar ),
    by = c("lon", "lat")
  )

# v3 Balland
df_bl_v3 <- df_bl %>% 
  left_join(
    nc_to_df("data/whc_mct_40y_v3.nc", "whc_mct_40y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_40y = myvar ),
    by = c("lon", "lat")
  )  %>% 
  left_join(
    nc_to_df("data/whc_mct_20y_v3.nc", "whc_mct_20y", do_get_ilon_ilat=FALSE) %>% 
      drop_na() %>% 
      rename( whc_mct_20y = myvar ),
    by = c("lon", "lat")
  )
```

Calculate rooting depth.
```{r echo=FALSE}
# v1
df_40y_sr_v1 <- df_sr_v1 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_40y, whc_t, whc_s))
save(df_40y_sr_v1, file = "./data/df_40y_sr_v1.Rdata")

df_20y_sr_v1 <- df_sr_v1 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_20y, whc_t, whc_s))
save(df_20y_sr_v1, file = "./data/df_20y_sr_v1.Rdata")

df_40y_bl_v1 <- df_bl_v1 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_40y, whc_t, whc_s))
save(df_40y_bl_v1, file = "./data/df_40y_bl_v1.Rdata")

df_20y_bl_v1 <- df_bl_v1 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_20y, whc_t, whc_s))
save(df_20y_bl_v1, file = "./data/df_20y_bl_v1.Rdata")

# v3
df_40y_sr_v3 <- df_sr_v3 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_40y, whc_t, whc_s))
save(df_40y_sr_v3, file = "./data/df_40y_sr_v3.Rdata")

df_20y_sr_v3 <- df_sr_v3 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_20y, whc_t, whc_s))
save(df_20y_sr_v3, file = "./data/df_20y_sr_v3.Rdata")

df_40y_bl_v3 <- df_bl_v3 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_40y, whc_t, whc_s))
save(df_40y_bl_v3, file = "./data/df_40y_bl_v3.Rdata")

df_20y_bl_v3 <- df_bl_v3 %>% 
  rowwise() %>% 
  drop_na() %>% 
  mutate(zroot = calc_zroot(whc_mct_20y, whc_t, whc_s))
save(df_20y_bl_v3, file = "./data/df_20y_bl_v3.Rdata")
```

Distribution of rooting depth for v3 and different pedotransfer functions, 40 yr return period.
```{r echo=FALSE}
df_tmp <- df_20y_sr_v3 %>% 
  dplyr::select(zroot_sr = zroot) %>% 
  bind_cols(dplyr::select(df_20y_bl_v3, zroot_bl = zroot)) %>% 
  tidyr::gather(method, zroot, c(zroot_bl, zroot_sr))

df_tmp %>% 
  ggplot() +
  geom_histogram(
    aes(x = zroot/1000, y = ..density.., fill = method), 
    color = "black", alpha = 0.3, binwidth = 0.2, 
    position="identity") +
  xlim(0,20) +
  scale_fill_manual(name = "", values = c("black", "red"), labels = c("Balland", "SaxtonRawls")) +
  labs(title = "Distribution of rooting depth (m), v3")
#ggsave(paste0("fig/hist_zroot_saxtonrawls_balland_v3.pdf"), width=7, height=5)
```

Distribution of rooting depth for v3 and different fAPAR data (v1 = mean across 0.5 deg, v3 = max across 0.5 deg), 40 yr return period, both using Balland et al pedotransfer functions.
```{r echo=FALSE}
df_tmp <- df_20y_bl_v3 %>% 
  dplyr::select(lon, lat, v3 = zroot) %>% 
  left_join( dplyr::select(df_20y_bl_v1, v1 = zroot, lon, lat), by = c("lon", "lat")) %>% 
  tidyr::gather(method, zroot, c(v3, v1))

df_tmp %>% 
  ggplot() +
  geom_histogram(
    aes(x = zroot/1000, y = ..density.., fill = method), 
    color = "black", alpha = 0.3, binwidth = 0.2, 
    position="identity") +
  xlim(0,20) +
  scale_fill_manual(name = "", values = c("black", "red")) +
  labs(title = "Distribution of rooting depth (m), v3", x = "Rooting depth (m)")
#ggsave(paste0("fig/hist_zroot_saxtonrawls_balland_v3.pdf"), width=7, height=5)
```

Map of rooting depth, plot and write to file.
```{r echo=FALSE}
## v1
## 40 y
grid_zroot_40y_bl_v1 <- df_40y_bl_v1 %>% 
  df_to_grid("zroot", "halfdeg") 
grid_zroot_40y_sr_v1 <- df_40y_sr_v1 %>% 
  df_to_grid("zroot", "halfdeg") 

#plot_map(grid_zroot_40y_bl_v1/1000, lev = c(0, 10, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 20) # , file="fig/map_zroot_40y_bl_v1.pdf"

#plot_map(grid_zroot_40y_sr_v1/1000, lev = c(0, 10, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 20) # , file="fig/map_zroot_40y_sr_v1.pdf"

write_nc2(
  var = grid_zroot_40y_bl_v1/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_40y_bl_v1.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )
write_nc2(
  var = grid_zroot_40y_sr_v1/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_40y_sr_v1.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )

## 20 y
grid_zroot_20y_bl_v1 <- df_20y_bl_v1 %>% 
  df_to_grid("zroot", "halfdeg") 
grid_zroot_20y_sr_v1 <- df_20y_sr_v1 %>% 
  df_to_grid("zroot", "halfdeg") 

plot_map(grid_zroot_20y_bl_v1/1000, lev = c(0, 5, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 10, toplefttext = "zroot 20y v1") # , file="fig/map_zroot_20y_bl_v3.pdf"

#plot_map(grid_zroot_20y_sr_v1/1000, lev = c(0, 5, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 10) # , file="fig/map_zroot_20y_sr_v3.pdf"

write_nc2(
  var = grid_zroot_20y_bl_v1/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_20y_bl_v1.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )
write_nc2(
  var = grid_zroot_20y_sr_v1/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_20y_sr_v1.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )

## v3
## 40 y
grid_zroot_40y_bl_v3 <- df_40y_bl_v3 %>% 
  df_to_grid("zroot", "halfdeg") 
grid_zroot_40y_sr_v3 <- df_40y_sr_v3 %>% 
  df_to_grid("zroot", "halfdeg") 

# plot_map(grid_zroot_40y_bl_v3/1000, lev = c(0, 10, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 20) # , file="fig/map_zroot_40y_bl_v3.pdf"

# gg <- plot_map2(grid_zroot_40y_bl_v3/1000, nbin = 10, maxval = 30) # , file="fig/map_zroot_20y_bl_v3.pdf"
# gg + labs(title = expression(paste("z"[root], "(m)")))
# ggsave("fig/zroot_40y_v3.png", width = 10, height = 7)
# 
# plot_map(grid_zroot_40y_sr_v3/1000, lev = c(0, 10, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 20) # , file="fig/map_zroot_40y_sr_v3.pdf"

write_nc2(
  var = grid_zroot_40y_bl_v3/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_40y_bl_v3.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )
write_nc2(
  var = grid_zroot_40y_sr_v3/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_40y_sr_v3.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )

## 20 y
grid_zroot_20y_bl_v3 <- df_20y_bl_v3 %>% 
  df_to_grid("zroot", "halfdeg") 
grid_zroot_20y_sr_v3 <- df_20y_sr_v3 %>% 
  df_to_grid("zroot", "halfdeg") 

plot_map(grid_zroot_20y_bl_v3/1000, lev = c(0, 15, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 30, toplefttext = "zroot 20y v3") # , file="fig/map_zroot_20y_bl_v3.pdf"

# plot_map(grid_zroot_20y_sr_v3/1000, lev = c(0, 15, 10), color = c( "wheat", "tomato2", "tomato4", "darkorchid4" ), minval = 0, maxval = 30) # , file="fig/map_zroot_20y_sr_v3.pdf"

write_nc2(
  var = grid_zroot_20y_bl_v3/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_20y_bl_v3.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )
write_nc2(
  var = grid_zroot_20y_sr_v3/1000, 
  varnam = "zroot", 
  filnam_template = "/alphadata01/bstocker/sofun/output_nc_global_sofun/s1_fapar3g_v3_global.fland.nc", 
  outfilnam = "data/zroot_mct_20y_sr_v3.nc",
  lonnam_template = "lon", 
  latnam_template = "lat"
  )
```